### 2024-07-05 ms_figure_6_networks.Rmd
Rearrange Fig 6, add new panels

---
title: "Figure 6 - networks"
author: "Charlotte Soneson, Michael Stadler, Merle Skribbe"
date: "`r Sys.Date()`"
output: 
    html_document:
        theme: united
        toc: true
        toc_float: true
        code_folding: hide
params:
    rds150: "data/ipms_150_sce.rds"
    coords150: "data/visnetwork_coords_150_new.csv"
    rds500: "data/ipms_500_sce.rds"
    coords500: "data/visnetwork_coords_500_new.csv"
    rds150untag: "data/ipms_150_untag_sce.rds"
    interactors: "data/Gene_list_TFinteractors_Pombase20240226.txt"
    idmap: "data/id_mapping_table.txt"
    rad24_rad25_tbl: "data/FigS5_Rad24_Rad25_interactingTFs_list_motifs.txt"
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = c("png", "cairo_pdf"))
```

# Parameter values

```{r params}
params
```

# Load required packages and helper functions

```{r load-packages, class.source="fold-show"}
suppressPackageStartupMessages({
    library(visNetwork)
    library(SingleCellExperiment)
    library(dplyr)
    library(ggplot2)
    library(igraph)
    library(shiny)
    library(cowplot)
    library(ggplotify)
    library(ggraph)
    library(latex2exp)
    library(tidygraph)
    library(ggrepel)
    library(ComplexHeatmap)
    library(flextable)
    library(kableExtra)
    library(ggupset)
    library(tidyr)
})

source("params/plot_settings.R")
source("params/get_testres_function.R")
source("params/mapping_functions.R")
```

# Settings

```{r}
w_trad <-  84 # heatmap body width (mm)
h_trad <- 178 # total heatmap height (mm)

fs <- 6  # small font size
fl <- 10 # large font size
ft <- 12 # title font size
```

# Read data

Here we load `SingleCellExperiment` objects corresponding to the low- and 
high-salt IP-MS experiments, as well as an object corresponding to the low-salt
experiments, where each condition was compared to the untagged controls 
rather than the large complement group as in the other two objects.

```{r load-data, class.source="fold-show"}
sce150 <- readRDS(params$rds150)
sce500 <- readRDS(params$rds500)
sce150untag <- readRDS(params$rds150untag)
```

We also load a table with mappings between different types of feature 
identifiers, and a table with known, annotated physical interactions 
(obtained from PomBase).

```{r load-idmap, class.source="fold-show"}
idmap <- read.delim(params$idmap)

interactors <- read.delim(params$interactors) |>
    dplyr::select(Systematic.ID, Gene.name, Physical.interactors) |>
    tidyr::separate_rows(Physical.interactors, sep = ",") |>
    dplyr::mutate(Gene.name = .capitalize(Gene.name),
                  Physical.interactors = .capitalize(Physical.interactors))
```

# Helper function

```{r}
## Overwrite the get_legend function from cowplot temporarily, 
## since the cowplot one doesn't work with ggplot2 3.5.0
## see https://github.com/wilkelab/cowplot/issues/202

get_legend <- function(plot, legend = NULL) {
    gt <- ggplotGrob(plot)

    pattern <- "guide-box"
    if (!is.null(legend)) {
        pattern <- paste0(pattern, "-", legend)
    }

    indices <- grep(pattern, gt$layout$name)

    not_empty <- !vapply(
        gt$grobs[indices], 
        inherits, what = "zeroGrob", 
        FUN.VALUE = logical(1)
    )
    indices <- indices[not_empty]

    if (length(indices) > 0) {
        return(gt$grobs[[indices[1]]])
    }
    return(NULL)
}
```

# Calculate pairwise protein similarities

To define the similarity between two proteins, the following procedure is used: 
Let $t_{ij}$ and $padj_{ij}$ denote the t-statistic and adjusted p-value for 
protein $i$ in comparison $j$, respectively. Then define 
$$\tilde{t}_{ij} = \left\{\begin{array}{cl}
    0 & \quad \textrm{if } t_{ij} <= 0 \textrm{ or } padj_{ij} > 0.1 \\
    t_{ij} & \quad \textrm{otherwise}
  \end{array}\right.$$

The similarity between proteins $i$ and $k$ is then defined by 
$$sim_{ik} = \frac{\sum_{j=1}^N\tilde{t}_{ij}\tilde{t}_{kj}}{max\left(\sum_{j=1}^N\tilde{t}_{ij}, \sum_{j=1}^N\tilde{t}_{kj}\right)}$$

or, if written in another way:
$$sim_{ik} = min\left(\frac{\sum_{j=1}^N\tilde{t}_{ij}\tilde{t}_{kj}}{\sum_{j=1}^N\tilde{t}_{ij}},\frac{\sum_{j=1}^N\tilde{t}_{ij}\tilde{t}_{kj}}{\sum_{j=1}^N\tilde{t}_{kj}}\right)$$
it can be interpreted as the smallest of two weighted means.

The following helper function extracts t-statistics and adjusted p-values 
and calculates the truncated values mentioned above ($\tilde{t}$).

```{r, class.source="fold-show"}
## Extract t-statistics and adjusted p-values and threshold
getTruncT <- function(sce, adjpThr = 0.1) {
    ## Extract adjusted p-values and log-fold changes
    tc <- .getTestCols(sce)

    adjpvals <- tc$adjp
    tstats <- tc$tstat
    colnames(adjpvals) <- .getProteinNameFromComparison(colnames(adjpvals))
    colnames(tstats) <- .getProteinNameFromComparison(colnames(tstats))
    adjpvals <- as.matrix(adjpvals[, grep("untagged|Untagged", colnames(adjpvals), invert = TRUE)])
    tstats <- as.matrix(tstats[, grep("untagged|Untagged", colnames(tstats), invert = TRUE)])
    stopifnot(colnames(adjpvals) == colnames(tstats),
              rownames(adjpvals) == rownames(tstats))

    ## Need to pass a relatively loose significance threshold in at least one IP
    ## (don't want things to be similar just because they are weakly but consistently
    ## changing in many IPs)
    tstats[adjpvals > adjpThr] <- 0
    tstats[tstats <= 0] <- 0
    tstats[is.na(tstats)] <- 0

    tstats
}
```

We also define a function to calculate the similarities between each pair 
of proteins.

```{r, class.source="fold-show"}
## Helper function to get matrix of similarities
simsFromResults <- function(sce, adjpThr = 0.1) {

    tstats <- getTruncT(sce = sce, adjpThr = adjpThr)

    simil <- tstats %*% t(tstats)
    simil0 <- simil
    for (i in seq_len(nrow(simil))) {
        for (j in seq(i, ncol(simil))) {
            simil[i, j] <- simil0[i, j] / max(sum(tstats[i, ]), sum(tstats[j, ]))
            simil[j, i] <- simil0[j, i] / max(sum(tstats[i, ]), sum(tstats[j, ]))
        }
    }
    simil[is.na(simil)] <- 0

    ## Don't need self-connections
    diag(simil) <- 0

    stopifnot(rownames(simil) == colnames(simil))

    list(simil = simil, tstats = tstats, TFs = colnames(tstats))
}
```

To illustrate, let's calculate the similarity for a specific example pair 
(Atf1 and Pcr1):

```{r, class.source="fold-show"}
tmp <- getTruncT(sce = sce150, adjpThr = 0.1)[c("Atf1", "Pcr1"), ]
round(tmp[, colSums(tmp) != 0], 1)
```

$$sim_{\textrm{atf1,pcr1}} = \min\left(\frac{3\cdot2.3+2.4\cdot2.5+12.2\cdot7.2+8.4\cdot6.9}{2.3+2.5+7.2+6.9},\frac{3\cdot2.3+2.4\cdot2.5+12.2\cdot7.2+8.4\cdot6.9}{3+2.4+12.2+8.4}\right) = min\left(8.4,6.1\right) = 6.1$$

```{r, class.source="fold-show"}
simsFromResults(sce150[c("Atf1", "Pcr1"), ], adjpThr = 0.1)$simil
```

Next, we calculate similarities for all pairs of proteins in each of the 
data sets.

```{r get-similarities, class.source="fold-show"}
simil150 <- simsFromResults(sce150)
simil500 <- simsFromResults(sce500)
maxSimil <- round(max(max(simil150$simil, na.rm = TRUE),
                      max(simil500$simil, na.rm = TRUE)), 1)
```

# Generate networks

Given the matrix of similarities calculated above, we create a graph by first 
thresholding the similarities (setting all values below a given threshold to 0), 
and then considering the remaining (non-zero) values in the matrix as 
representing edges in the graph. Thus, two proteins in the graph are connected 
if their enrichment patterns across the IP experiments are similar. 
Only graph components with at least two proteins (i.e., no singletons) are 
displayed.

```{r make-networks}
createNetwork <- function(similList, threshold, known_interactors, settings = "igraph") {
    ## Get similarity matrix and apply threshold
    simil <- similList$simil
    tfs <- similList$TFs
    simil[simil < threshold] <- 0
    simil[is.na(simil)] <- 0

    ## Create graph
    gr <- graph_from_adjacency_matrix(simil, weighted = TRUE)
    gr <- as.undirected(gr, mode = "collapse", edge.attr.comb = list(weight = "mean"))

    ## Remove singletons
    comps <- components(gr)
    too_small <- which(table(comps$membership) < 2)
    keep <- V(gr)[!(comps$membership %in% too_small)]
    gr <- induced_subgraph(gr, keep)

    ## Set attributes for vertices and edges
    ## Need col2hex here since visNetwork doesn't understand all R colors (e.g. gray70)
    V(gr)$color <- gplots::col2hex(bg_color)
    V(gr)$color[names(V(gr)) %in% tfs] <- main_colors[5]
    V(gr)$vertexclass <- "other"
    V(gr)$vertexclass[names(V(gr)) %in% tfs] <- "Bait"

    if (settings == "visNetwork") {
        V(gr)$shape <- "triangle"
        V(gr)$size <- 15
        V(gr)$label.cex = 2.08
    } else if (settings == "igraph") {
        V(gr)$shape <- "circle"
    } else {
        stop("Unknown 'settings' parameter")
    }

    V(gr)$shape[names(V(gr)) %in% tfs] <- "square"
    E(gr)$value <- E(gr)$weight
    E(gr)$width <- E(gr)$weight / max(simil) * 1.5

    ## Indicate annotated physical interactions
    eL <- as_edgelist(gr)
    for (i in seq_along(E(gr))) {
        v1 <- eL[i, 1]
        v2 <- eL[i, 2]
        tmp <- known_interactors |>
            dplyr::filter(((Systematic.ID == v1 | Gene.name == v1) & 
                               Physical.interactors == v2) | 
                              ((Systematic.ID == v2 | Gene.name == v2) & 
                                   Physical.interactors == v1))
        if (nrow(tmp) == 0) {
            ## Novel interaction
            ## One of the elements is a transcription factor -> color orange
            if (v1 %in% tfs || v2 %in% tfs) {
                E(gr)[i]$color <- main_colors[3]
                E(gr)[i]$edgeclass <- "Unannotated TF interaction"
            } else{
                E(gr)[i]$color <- gplots::col2hex(na_color)
                E(gr)[i]$edgeclass <- "Other"
            }
        } else {
            E(gr)[i]$color <- gplots::col2hex(na_color)
            E(gr)[i]$edgeclass <- "Other"
        }
    }
    E(gr)$edgeclass <- factor(E(gr)$edgeclass, 
                              levels = c("Unannotated TF interaction", 
                                         "Other"))
    gr
}
```

```{r}
#| echo: false
#| eval: false

## Code for shiny app to compare thresholds and generate coordinate csv file
shinyAppThresholds <- function(similList, known_interactors) {
    server <- function(input, output) {
        output$visnetwork <- renderVisNetwork({
            gr <- createNetwork(similList, threshold = input$threshold, 
                                known_interactors = known_interactors, 
                                settings = "visNetwork")
            visNetwork::visIgraph(gr)
        })

        # Get position info
        observeEvent(input$store_position, {
            visNetworkProxy("visnetwork") |> visGetPositions()
        })

        output$coordinates <- renderPrint({
            nodes_positions()
        })

        ## Format positions
        nodes_positions <- reactive({
            positions <- input$visnetwork_positions
            if (!is.null(positions)){
                nodes_positions <- do.call("rbind", lapply(positions, function(x){data.frame(x = x$x, y = x$y)}))
                nodes_positions$id <- names(positions)
                nodes_positions
            } else {
                NULL
            }
        })

        output$downloadCsv <- shiny::downloadHandler(
            filename = function() {
                "graph_coordinates.csv"
            },
            content = function(file) {
                utils::write.table(nodes_positions(), file,
                                   row.names = FALSE, sep = ",",
                                   quote = FALSE)
            }
        )
    }

    ui <- fluidPage(
        sliderInput("threshold", label = "Score threshold", min = 0, max = maxSimil,
                    value = 6, step = 0.01),
        visNetworkOutput("visnetwork", width = "100%", height = "1000px"),
        verbatimTextOutput("coordinates"),
        actionButton("store_position", "Store positions"),
        downloadButton("downloadCsv", "Download")
    )

    shinyApp(ui = ui, server = server)
}
shiny::runApp(shinyAppThresholds(simil150, interactors))
shiny::runApp(shinyAppThresholds(simil500, interactors))
```

Next, we create networks for the low- and high-salt experiments.

```{r}
gr150 <- createNetwork(similList = simil150, threshold = 6, 
                       known_interactors = interactors, settings = "igraph")
gr500 <- createNetwork(similList = simil500, threshold = 6, 
                       known_interactors = interactors, settings = "igraph")
```

# 14-3-3 protein interactions

We extract all IPs where either `Rad24` or `Rad25` are pulled down, 
using a relatively lenient adjusted p-value cutoff.

```{r rad24-rad25}
#| fig.width: 9
#| fig.height: 15

## Get all t-statistics and visualize
res150untag <- .getTestCols(sce150untag, adjp_cutoff = 0.01, logfc_cutoff = 1)
rad <- res150untag$tstat[c("Rad24", "Rad25"), ]
colnames(rad) <- .getProteinNameFromComparison(colnames(rad))
rad <- t(rad)
hmrad <- Heatmap(rad, 
                 cluster_rows = TRUE, 
                 cluster_columns = FALSE, 
                 col = circlize::colorRamp2(
                     breaks = seq(1.5, 13.0, length.out = 64),
                     colors = colorRampPalette(
                         binary_heatmap_colors[c("FALSE", "TRUE")])(64)),
                 border = TRUE, 
                 border_gp = gpar(lwd = 0.5),
                 name = "Mod.\nt-stat.",
                 na_col = binary_heatmap_colors["FALSE"],
                 column_names_gp = gpar(fontsize = fl),
                 row_names_gp = gpar(fontsize = fs),
                 show_row_names = TRUE, 
                 show_column_names = TRUE, 
                 use_raster = FALSE,
                 show_heatmap_legend = TRUE, 
                 width = unit(w_trad / 1.5, "mm"), 
                 heatmap_height = unit(1.4 * h_trad, "mm"),
                 column_title_gp = gpar(fontsize = ft),
                 heatmap_legend_param = list(
                     title_gp = gpar(fontsize = fl),
                     border = "gray10",
                     border_gp = gpar(lwd = 0.5)
                 ))
set.seed(42L)
hm_rad <- grid.grabExpr(
    hmrad <- draw(hmrad, merge_legend = TRUE,
                 heatmap_legend_side = "right",
                 annotation_legend_side = "right"),
    width = w_trad / 1.5, height = 1.4 * h_trad
)

## List of experiments where rad24 and/or rad25 were pulled down
comps <- colnames(res150untag$interactor)[colSums(
    res150untag$interactor[c("Rad24", "Rad25"), ], na.rm = TRUE) > 0]
explist <- .getUniProtIdFromComparison(comps, idmap = idmap)
cat(explist)
gnames <- .getProteinNameFromComparison(comps)
cat(gnames)

## Read table
set_flextable_defaults(font.size = 8)
rad24_rad25_tbl <- read.delim(params$rad24_rad25_tbl)[, 1:2] |>
    setNames(c("TF", "R-x-x-[S/T]-x-P")) |>
    dplyr::mutate(TF = .capitalize(TF))
stopifnot(all(gnames %in% rad24_rad25_tbl$TF))
ft <- flextable(rad24_rad25_tbl) |>
  width(width =  c(0.5, 1.25)) |>
  set_table_properties(layout = "fixed")
ggtbl <- ggplot() +
    theme_void() +
    annotation_custom(grid::rasterGrob(as_raster(ft)),
                      xmin = -Inf, xmax = Inf, ymin = 0, ymax = 1)

data.frame(`Number of TFs interacting with at least one 14-3-3 protein` = 
               nrow(rad24_rad25_tbl),
           `Number of such TFs with at least one optimal 14-3-3 binding motif` = 
               sum(rad24_rad25_tbl$`R-x-x-[S/T]-x-P` != ""),
           check.names = FALSE) |>
    pivot_longer(cols = everything()) |>
    kbl(col.names = NULL) |>
    kable_styling()

# UpSet plot
dfupset <- as.data.frame(t(res150untag$interactor[c("Rad24", "Rad25"), ])) |>
    tibble::rownames_to_column("TF") |>
    dplyr::mutate(TF = .getProteinNameFromComparison(TF))

ovL <- as.list(as.data.frame(dfupset[, c("Rad24", "Rad25")]))
ovL <- lapply(ovL, function(i) dfupset$TF[which(i)])

pd <- stack(ovL) |>
    group_by(values) |>
    summarise(ind = list(ind)) |>
    dplyr::left_join(rad24_rad25_tbl, by = join_by("values" == "TF")) |>
    dplyr::mutate(has_motif = !is.na(`R-x-x-[S/T]-x-P`) & `R-x-x-[S/T]-x-P` != "")

ggrad_upset <- ggplot(pd, aes(x = ind)) +
    geom_bar(aes(fill = has_motif)) +
    scale_fill_manual(values = c(`TRUE` = main_colors[3], `FALSE` = main_colors[5])) +
    scale_x_upset() +
    labs(x = element_blank(), y = "Number of IPs", fill = "Has motif: ") +
    theme_cowplot(12) +
    theme(axis.ticks.x = element_blank(),
          legend.position = "bottom") +
    theme_combmatrix(combmatrix.label.extra_spacing = 5,
                     combmatrix.label.make_space = TRUE,
                     combmatrix.label.width = unit(18, "mm")) +
    guides(fill = guide_legend(nrow = 1))
```

# Figure 6

```{r}
#| fig.width: 12
#| fig.height: 5

## Create small example similarity matrix and graph, for three selected proteins
poi <- c("Atf1", "Pcr1", "Arp8")
simil150example <- simsFromResults(sce150[poi, ])
gr150example <- createNetwork(similList = simil150example, threshold = 4.5,
                              known_interactors = interactors, 
                              settings = "igraph")

## General formula
## See https://uliniemann.com/blog/2022-02-21-math-annotations-in-ggplot2-with-latex2exp/
eqn1 <- data.frame(x = 0, y = 0, 
                   label = r"( $sim_{ik} = min\left(\frac{\sum_{j=1}^N\tilde{t}_{ij}\tilde{t}_{kj}}{\sum_{j=1}^N\tilde{t}_{ij}},\frac{\sum_{j=1}^N\tilde{t}_{ij}\tilde{t}_{kj}}{\sum_{j=1}^N\tilde{t}_{kj}}\right)$ )")

ggeqn1 <- ggplot(eqn1) +
    geom_text(aes(x = x, y = y, 
                  label = TeX(label, output = "character"), 
                  color = label), 
              parse = TRUE,
              key_glyph = draw_key_blank,
              size = 16/.pt) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_reverse(expand = c(0, 0)) + 
    scale_color_manual(values = "transparent", labels = TeX(eqn1$label,
                                                            italic = TRUE), 
                       name = "") + 
    guides(color = guide_legend(label.position = "top")) + 
    theme_void() + 
    theme(legend.text = element_text(size = 11))

## Specific example (atf1, pcr1)
eqn2 <- data.frame(x = 0, y = 0, 
                   label = r"( $sim_{Atf1,Pcr1} = min\left(\frac{3 \cdot 2.3 + 2.4 \cdot 2.5 + 12.2 \cdot 7.2 + 8.4 \cdot 6.9}{2.3 + 2.5 + 7.2 + 6.9},\frac{3 \cdot 2.3 + 2.4 \cdot 2.5 + 12.2 \cdot 7.2 + 8.4 \cdot 6.9}{3 + 2.4 + 12.2 + 8.4}\right) = min\left(8.4, 6.1\right) = 6.1$ )")
# eqn2 <- data.frame(x = 0, y = 0, 
#                    label = r"( $sim_{\textrm{Atf1,Pcr1}} = \min\left(\frac{3 \cdot 2.3 + 2.4 \cdot 2.5 + 12.2 \cdot 7.2 + 8.4 \cdot 6.9}{2.3 + 2.5 + 7.2 + 6.9},\frac{3 \cdot 2.3 + 2.4 \cdot 2.5 + 12.2 \cdot 7.2 + 8.4 \cdot 6.9}{3 + 2.4 + 12.2 + 8.4}\right) = min\left(8.4, 6.1\right) = 6.1$ )")

ggeqn2 <- ggplot(eqn2) +
    geom_text(aes(x = x, y = y, 
                  label = TeX(label, output = "character")), 
              parse = TRUE, 
              size = 11/.pt) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_reverse(expand = c(0, 0)) + 
    theme_void()

## Truncated t-statistics
tstats_vis <- data.frame(simil150example$tstats) |>
    tibble::rownames_to_column("Protein") |>
    tidyr::pivot_longer(names_to = "IP", values_to = "Value", -Protein) |>
    dplyr::mutate(Protein = factor(Protein, levels = poi))

## Reorder IPs for better visualization
enr <- unique(tstats_vis$IP[tstats_vis$Value > 0])
notenr <- setdiff(tstats_vis$IP, enr)
allips <- c(notenr, enr)
intlen <- round(length(notenr) / (length(enr) + 1))
idx <- seq(from = intlen, by = intlen, length.out = length(enr))
levs <- allips[order(c(seq_along(notenr), idx + 0.5))]
tstats_vis$IP <- factor(tstats_vis$IP, levels = levs)

## Plot truncated tstats across all IPs and label relevant points
labels <- tstats_vis |>
    dplyr::filter(Value > 0) |>
    dplyr::pull(IP) |>
    unique()

tplot <- ggplot(tstats_vis, aes(x = IP, y = Value, color = Protein, fill = Protein)) + 
        geom_col() + 
        geom_text(
            aes(label = ifelse(IP %in% labels & Value > 0, round(Value, digits = 1), "")),
            fontface ="plain", color = "black", size = 2.5, nudge_y = 2) +
        scale_color_manual(values = c(main_colors[1], main_colors[5], "black")) +
        scale_fill_manual(values = c(main_colors[1], main_colors[5], "black")) +
        scale_y_continuous(expand = c(0, 1)) + 
        facet_wrap(~ Protein, ncol = 1) +
        labs(x = "IP", y = "Thresholded mod. t-stat.") + 
        theme_cowplot(font_size = 10) + 
        theme(legend.position = "none", 
              axis.text.x = element_blank(), 
              axis.ticks.x = element_blank(), 
              strip.background = element_rect(fill = NA, colour = NA), 
              strip.text = element_text(hjust = 0, size = 10, face = "bold"))

## Similarity matrix
similsub <- simil150example$simil[poi, poi] |>
    as.data.frame() |>
    tibble::rownames_to_column("row") |>
    tidyr::pivot_longer(names_to = "column", values_to = "Similarity", -row) |>
    dplyr::mutate(Similarity = round(Similarity, digits = 1),
                  row = factor(row, levels = rev(poi)),
                  column = factor(column, levels = poi)) |>
    dplyr::mutate(simlab = ifelse(Similarity == 0, "", Similarity)) |>
    dplyr::mutate(Similarity = ifelse(Similarity == 0, NA, Similarity))
similplot <- ggplot(similsub, aes(x = column, y = row, fill = Similarity)) + 
    geom_tile() + 
    geom_text(aes(label = simlab, color = Similarity > 4), 
              fontface = "bold") +
    scale_color_manual(values = c(`TRUE` = "white", `FALSE` = "black")) + 
    scale_fill_gradient(low = "white", high = binary_heatmap_colors["TRUE"], 
                    na.value = na_color, 
                    limits = range(similsub$Similarity, na.rm = TRUE)) + 
    theme_cowplot() + 
    theme(axis.line = element_blank(), 
          axis.ticks = element_blank(),
          legend.position = "bottom",
          legend.title = element_text(size = 10, vjust = 0.82),
          legend.margin = margin(0, 0, 0, 0),
          legend.box.spacing = unit(0, "pt")) + 
    labs(x = "", y = "") + 
    guides(color = "none")

## Induced graph
ggr150example <- ggraph(as_tbl_graph(gr150example), 
                        layout = cbind(x = c(-300, 800), y = c(800, -300)), 
                        weights = weight) + 
    geom_edge_link(aes(color = edgeclass, width = weight)) +
    scale_edge_color_manual(values = c(`Unannotated TF interaction` = main_colors[3], 
                                       Other = gplots::col2hex(na_color)),
                            name = "") + 
    scale_edge_width(range = c(1.2, 1.9), guide = "none") + 
    geom_node_point(size = 4, aes(fill = vertexclass, shape = vertexclass), 
                    color = "black", stroke = 0.1) + 
    scale_shape_manual(values = c(other = 21, Bait = 22), name = "") + 
    scale_fill_manual(values = c(other = gplots::col2hex(bg_color), Bait = main_colors[5]),
                      name = "") + 
    geom_node_text(aes(label = name), size = 3.5, repel = TRUE, 
                   nudge_x = 0, nudge_y = -5) + 
    theme_graph(base_family = "sans", plot_margin = margin(30, 10, 40, 10)) + 
    guides(fill = guide_legend(override.aes = list(size = 4, alpha = 1))) + 
    coord_cartesian(clip = "off")

## Put together
fig5a <- cowplot::plot_grid(
    cowplot::plot_grid(
        tplot, 
        get_legend(ggeqn1, "right"), 
        similplot + theme(legend.position = "bottom"), 
        ggr150example + theme(legend.position = "none"), 
        align = "v", axis = "t", 
        labels = c("A", "", "", ""), nrow = 1, 
        rel_widths = c(1.1, 0.9, 1, 0.5)
    ),
    ggeqn2, ncol = 1, labels = "", 
    rel_heights = c(3, 1)
)
```

## Put together

```{r Fig6}
#| fig.width: 10
#| fig.height: 10

fig6c <- ggdraw() + 
    draw_image("schematics/Fig6C_Pho7_protein_domains_cartoon_Arial_RGB.png")

coords150 <- read.csv(params$coords150)
coords500 <- read.csv(params$coords500)
coords500$y <- -coords500$y

.plotGraph <- function(grph, coords, nudge_y) {
    ggraph(as_tbl_graph(grph), 
                 layout = coords, 
                 weights = weight) + 
    geom_edge_link(aes(color = edgeclass, width = weight)) +
    scale_edge_color_manual(values = c(`Unannotated TF interaction` = main_colors[3], 
                                       Other = gplots::col2hex(na_color)),
                            breaks = "Unannotated TF interaction",
                            name = "") + 
    scale_edge_width(range = c(0.2, 0.9), guide = "none") + 
    geom_node_point(size = 2, aes(fill = vertexclass, shape = vertexclass), 
                    color = "black", stroke = 0.1) + 
    scale_shape_manual(values = c(other = 21, Bait = 22), name = "",
                       breaks = "Bait") + 
    scale_fill_manual(values = c(other = gplots::col2hex(bg_color), Bait = main_colors[5]),
                      name = "", breaks = "Bait") + 
    geom_node_text(aes(label = name), size = 2.5, repel = FALSE, 
                   nudge_x = 0, nudge_y = nudge_y) + 
    theme_graph(base_family = "sans", plot_margin = margin(10, 10, 10, 10)) + 
    guides(fill = guide_legend(override.aes = list(size = 4, alpha = 1)))
}
ggr150 <- .plotGraph(grph = gr150, coords = as.matrix(coords150[, c("x", "y")]),
                     nudge_y = -45)
ggr500 <- .plotGraph(grph = gr500, coords = as.matrix(coords500[, c("x", "y")]),
                     nudge_y = -35)

cowplot::plot_grid(
    cowplot::plot_grid(
        cowplot::plot_grid(
            ggr150 + theme(legend.position = "none"), 
            ggr500 + theme(legend.position = "none"), 
            nrow = 1, labels = c("A", "B")
        ),
        get_legend(ggr150 + theme(legend.position = "bottom"), "bottom"),
        nrow = 2, labels = c("", ""), rel_heights = c(1, 0.25)
    ),
    cowplot::plot_grid(fig6c, ggrad_upset,
                       nrow = 1, labels = c("C", "D")), 
    rel_heights = c(5, 2), 
    ncol = 1
)
```

## Supplementary figure

We extract all IPs where either `Rad24` or `Rad25` are pulled down, 
using a relatively lenient adjusted p-value cutoff.

```{r suppFig6}
#| fig.width: 9
#| fig.height: 15

plot_grid(
    fig5a, 
    plot_grid(hm_rad, ggtbl, nrow = 1, align = "h", axis = "b", 
              rel_widths = c(2, 2.5), labels = c("B", "C")),
    rel_heights = c(1, 3), ncol = 1,
    labels = c("A", "")
)
```

# Session info

<details>
<summary><b>
Session info
</b></summary>
```{r}
sessionInfo()
```
</details>