### 2024-07-09 ms_figure_7_heterodimer.Rmd
Add genome logfc plot (Fig 7)

---
title: "Figure 7 - heterodimer"
author: "Charlotte Soneson, Michael Stadler, Merle Skribbe"
date: "`r Sys.Date()`"
output: 
    html_document:
        theme: united
        toc: true
        toc_float: true
        code_folding: hide
params:
    rds500: "data/ipms_500_sce.rds"
    rdsrnaseq_del: "data/rnaseq_del.rds"
    rdsrnaseq_ko: "data/rnaseq_ko.rds"
    dist_from_ne: "data/Fig7_lacR_quantification_distance_radius.txt"
    gtf: "reference/Schizosaccharomyces_pombe.ASM294v2.55.gtf"
    idmap: "data/id_mapping_table.txt"
    rnatracks_ko: "data/genome_browser_tracks_rnaseq_ko.csv.gz"
    chiptracks_ko: "data/genome_browser_tracks_chipseq_ko.csv.gz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = c("png", "cairo_pdf"))
```

# Parameter values

```{r params}
params
```

# Load required packages and helper functions

```{r load-packages, class.source="fold-show"}
suppressPackageStartupMessages({
    library(SingleCellExperiment)
    library(ggplot2)
    library(cowplot)
    library(dplyr)
    library(ggrepel)
    library(colorspace)
    library(DESeq2)
    library(ggh4x)
    library(ggtext)
    library(patchwork)
    library(ggsignif)
    library(kableExtra)
    library(scales)
    library(broom)
})

source("params/plot_settings.R")
source("params/get_testres_function.R")
source("params/mapping_functions.R")
```

# Read data

We load a `SingleCellExperiment` object, containing data and results from the 
high-salt IP-MS experiment.

```{r load-data, class.source="fold-show"}
sce500 <- readRDS(params$rds500)
```

Next, we load two `SummarizedExperiment` objects with quantifications 
from RNA-seq experiments.

```{r, class.source="fold-show"}
rnaseq_del <- readRDS(params$rdsrnaseq_del)
rnaseq_ko <- readRDS(params$rdsrnaseq_ko)
idmap <- read.delim(params$idmap)
```

# Figure 7

## Zone quantification

```{r}
## Read data with distances from nuclear envelope + cell radius and 
## calculate relative distance from NE
dist_ne <- read.delim(params$dist_from_ne) |>
    group_by(Condition, Replicate) |>
    mutate(rel_distance_from_ne = Distance_from_nuc_envelope / Radius_of_cell) |>
    mutate(Condition = ifelse(Condition == "ntu1D", "ntu1\U141E", 
                              ifelse(Condition == "ntu2D", "ntu2\U141E", Condition))) |>
    mutate(SampleID = factor(paste0(Condition, "\n", Replicate),
                             levels = c("wt\nrep1", "wt\nrep2", "ntu1\U141E\nrep1",
                                        "ntu1\U141E\nrep2", "ntu2\U141E\nrep1",
                                        "ntu2\U141E\nrep2")))

## Remove one cell where Distance was slightly larger than the Radius
# table(dist_ne$rel_distance_from_ne > 1)
dist_ne <- dist_ne |>
    filter(rel_distance_from_ne >= 0 & rel_distance_from_ne <= 1) |>
    dplyr::mutate(Condition = factor(Condition, 
                                     levels = c("wt", "ntu1\U141E", "ntu2\U141E")),
                  Replicate = factor(Replicate, 
                                     levels = c("rep1", "rep2")))

## Get zone classification
zone_thresholds <- 1 - c(1, sqrt(2), sqrt(3)) / sqrt(3)
zone_byrep <- dist_ne |>
    mutate(Zone = ifelse(rel_distance_from_ne > zone_thresholds[1], "Zone III", 
                         ifelse(rel_distance_from_ne > zone_thresholds[2], 
                                "Zone II", "Zone I"))) |>
    group_by(Condition, Replicate, Zone) |>
    tally() |>
    group_by(Condition, Replicate) |>
    mutate(fraction_by_zone = n / sum(n)) |>
    mutate(Zone = factor(Zone, levels = c("Zone I", "Zone II", "Zone III"))) |>
    ungroup()
zone_byrep |>
    mutate(fraction_by_zone = label_percent(accuracy = 0.01)(fraction_by_zone)) |>
    rename(`Number of cells` = n, 
           `Percentage of cells` = fraction_by_zone) |>
    kbl() |>
    kable_styling()

## Average zone distribution over replicates
zone_avg <- zone_byrep |> 
    dplyr::group_by(Zone, Condition) |> 
    dplyr::summarize(avg_fraction_by_zone = mean(fraction_by_zone), 
                     sd_fraction_by_zone = sd(fraction_by_zone), 
                     .groups = "drop")
```

```{r}
## Fraction of cells in Zone I
zone_byrep |> filter(Zone == "Zone I") |>
    group_by(Zone, Condition) |>
    summarize(average_percentage = label_percent(accuracy = 0.01)(mean(fraction_by_zone)), 
              individual_percentages = paste(label_percent(accuracy = 0.01)(fraction_by_zone), 
                                             collapse = ","), 
              .groups = "drop") |>
    rename(`Average percentage` = average_percentage,
           `Percentages for replicates` = individual_percentages) |>
    kbl() |>
    kable_styling()

## Fraction of cells in Zone III
zone_byrep |> filter(Zone == "Zone III") |>
    group_by(Zone, Condition) |>
    summarize(average_percentage = label_percent(accuracy = 0.01)(mean(fraction_by_zone)), 
              individual_percentages = paste(label_percent(accuracy = 0.01)(fraction_by_zone), 
                                             collapse = ","), 
              .groups = "drop") |>
    rename(`Average percentage` = average_percentage,
           `Percentages for replicates` = individual_percentages) |>
    kbl() |>
    kable_styling()
```

### Plot relative distance to perimeter in all conditions

```{r}
#| warning: false

## Number of cells per condition
table(dist_ne$Replicate, dist_ne$Condition)

get_xlabs <- function(x) {
    vapply(x, function(xx) sub("\n", "<br>", paste0("*", xx, "*", "<br>(n = ", 
                                                    length(which(paste0(dist_ne$Condition, "\n", 
                                                                        dist_ne$Replicate) == xx)), ")"), fixed = TRUE), "")
}

## Calculate p-values (average values per sample followed by linear model)
dps <- dist_ne |>
    group_by(SampleID, Condition, Replicate) |>
    summarize(mean_dist = mean(rel_distance_from_ne),
              median_dist = median(rel_distance_from_ne),
              .groups = "drop")
(aggreg <- summary(lm(median_dist ~ Condition, data = dps)) |> tidy())

## wt vs ntu1D
p_ntu1 <- paste0("p==", signif(aggreg$p.value[aggreg$term == "Conditionntu1\U141E"], 3))

## wt vs ntu2D
p_ntu2 <- paste0("p==", signif(aggreg$p.value[aggreg$term == "Conditionntu2\U141E"], 3))

(p_rel_dis <- ggplot(dist_ne, 
                     aes(x = SampleID, y = rel_distance_from_ne, 
                         fill = Condition)) + 
        geom_violin(trim = TRUE, bounds = c(0, 1)) +
        geom_signif(annotations = c(p_ntu1, p_ntu2),
                    xmin = c(1.5, 1.5), xmax = c(3.5, 5.5), 
                    y = c(1.05, 1.15),
                    tip_length = 0, step_increase = 0.1, parse = TRUE) + 
        geom_boxplot(width = 0.1, fill = "white") +
        labs(title = "Relative distance from nuclear envelope",
             x = "", y = "Distance from NE relative to radius") + 
        scale_fill_manual(values = c(na_color, main_colors[5], main_colors[1]),
                          name = "") + 
        scale_x_discrete(labels = get_xlabs) +
        scale_y_continuous(breaks = c(0, 0.5, 1), limits = c(0, 1.2)) + 
        theme_cowplot() + 
        theme(axis.text.x = element_markdown(),
              legend.text = element_text(face = "italic")))
```

### Plot zone distribution by condition

```{r}
#| warning: false

get_xlabs_bar <- function(x) {
    vapply(x, function(xx) sub("\n", "<br>", paste0("*", xx, "*", "<br>(n = ", 
                                                    length(which(dist_ne$Condition == xx)), ")"), fixed = TRUE), "")
}

(p_zone <- ggplot(zone_avg, aes(fill = Zone, y = 100 * avg_fraction_by_zone, 
                                x = Condition)) + 
     geom_bar(position = "stack", stat = "identity") +
     geom_text(aes(label = paste0(round(100 * avg_fraction_by_zone, 2), "%")), 
                   position = position_stack(vjust = 0.5), size = 4) +
     labs(title = "Relative zone distribution incl. both replicates", 
          x = "", y = "Relative distribution [%]") +
     scale_fill_manual(values = c(main_colors[1], main_colors[3], main_colors[5]),
                       name = "") + 
     scale_x_discrete(labels = get_xlabs_bar) + 
     theme_cowplot() + 
     theme(axis.text.x = element_markdown()))
```

## Ntu1/Ntu2 500 mM NaCl IP-MS

Here we extract and visualize the t-statistics for all proteins in the 
ntu1 and ntu2 high-salt IP-MS experiments. In each case the condition of 
interest is compared to a broad complement group consisting of all other 
high-salt IP-MS samples.

```{r tstat-500-ntu1-ntu2}
#| fig.width: 4
#| fig.height: 4
#| warning: false

## Extract t-statistics for the two comparisons
rdhigh <- rowData(sce500)
contrasts_high <- c(
    "Ntu1_500_plate_vs_compl_Ntu1_500_plate.t", 
    "Ntu2_500_plate_vs_compl_Ntu2_500_plate.t")
high <- as.data.frame(rdhigh[, c("einprotId", contrasts_high)]) |>
    setNames(c("einprotId", c("Ntu1", "Ntu2")))

## Plot t-statistics
(gg_ipdimer <- ggplot(high, aes(x = Ntu1, y = Ntu2)) +
    geom_abline(data = high |> 
                    dplyr::filter(einprotId %in% c("Ntu1", "Ntu2")), 
                aes(slope = mean(Ntu2 / Ntu1), intercept = 0), 
                linetype = "dashed", color = "grey50") + 
    geom_text_repel(data = high |> 
                        dplyr::filter(einprotId %in% c("Ntu1", "Ntu2") | 
                                          Ntu2 > 5),
                    aes(label = einprotId), min.segment.length = 0) + 
    geom_point(alpha = 0.25) + 
    geom_point(data = high |> 
                   dplyr::filter(einprotId %in% c("Ntu1", "Ntu2")) |>
                   dplyr::mutate(type = "Baits"),
               mapping = aes(fill = type), color = main_colors[3],
               shape = 21, alpha = 1, size = 2) + 
    scale_fill_manual(values = c(Baits = main_colors[3])) + 
    theme_cowplot() + 
    labs(x = "500 mM NaCl mod. t-stat. (Ntu1)", 
         y = "500 mM NaCl mod. t-stat. (Ntu2)") + 
    theme(strip.text = element_blank(),
          legend.position = "inside",
           legend.position.inside = c(0.8, 0.1)) + 
    guides(fill = guide_legend(nrow = 1, byrow = TRUE, title = "")))
```

## RNA-seq expression levels for genes around the tna1 locus

```{r}
#| fig.width: 8
#| fig.height: 6

wt_tpm <- assay(rnaseq_del[, rnaseq_del$Group_name == "wt_WT_mRNA"], "abundance")
genepos <- rowRanges(rnaseq_del)
mcols(genepos) <- NULL
genepos <- as.data.frame(genepos)
stopifnot(rownames(wt_tpm) == rownames(genepos))
genepos$ave_tpm <- rowMeans(wt_tpm)
genepos$id <- rownames(genepos)

chrom <- genepos["tna1", "seqnames"]
startpos <- genepos["tna1", "start"]
endpos <- genepos["tna1", "end"]
windowsize <- 30000

(genomeplot_tpm <- 
        ggplot(genepos |> 
                   dplyr::filter(seqnames == chrom & 
                                     start > startpos - windowsize & 
                                     end < endpos + windowsize), 
               aes(x = start / 1e6, y = ave_tpm, label = id)) +
        geom_hline(yintercept = 0, color = na_color) + 
        geom_point(size = 3, alpha = 0.8, aes(color = id == "tna1")) +
        scale_color_manual(values = c(`TRUE` = main_colors[4], `FALSE` = pt_color)) + 
        geom_text_repel(min.segment.length = 0, max.overlaps = Inf, size = 3, 
                        force = 3) + 
        labs(x = "Genome position (Mb)", y = "Gene expression (TPM)") + 
        scale_y_continuous(expand = expansion(mult = c(0, 0), add = c(200, 50))) + 
        theme_cowplot() + 
        theme(legend.position = "none"))
```

## RNA-seq differential expression analysis - KO experiments

```{r rnaseq-de}
#| fig.width: 15
#| fig.height: 4
#| message: false
#| warning: false

## Filter lowly expressed genes and create DESeqDataSet
rnaseqfilt_ko <- rnaseq_ko[rowSums(assay(rnaseq_ko, "counts") >= 5) >= 2, ]
dds <- DESeq2::DESeqDataSet(rnaseqfilt_ko, design = ~ MappedGene)
dds <- DESeq2::estimateSizeFactors(dds)

## Filter genes with high inferential variance
highInfRV <- 10^(-0.75)
dds <- dds[rowData(dds)$meanInfRV < highInfRV, ]

## Perform differential expression analysis
dds <- DESeq2::DESeq(dds)
groups <- setdiff(dds$MappedGene, "wt")
names(groups) <- groups
resList <- lapply(groups, function(gr) {
    DESeq2::results(dds, contrast = c("MappedGene", gr, "wt"), 
                    alpha = 0.05)
})

## Create MA plots
rngY <- range(sapply(resList, function(rl) rl$log2FoldChange))
rngX <- range(sapply(resList, function(rl) log10(rl$baseMean)))
maplots <- lapply(structure(names(resList), names = names(resList)), function(nm) {
    ggplot(as.data.frame(resList[[nm]]), 
           aes(x = log10(baseMean), y = log2FoldChange)) + 
        geom_point(mapping = aes(color = abs(log2FoldChange) > 3 & 
                                     padj < 0.001), alpha = 0.25) + 
        scale_color_manual(values = c(`TRUE` = main_colors[4], `FALSE` = pt_color)) + 
        coord_cartesian(xlim = rngX, ylim = rngY) + 
        geom_text_repel(data = as.data.frame(resList[[nm]]) |>
                            tibble::rownames_to_column("gene") |>
                            dplyr::filter(gene == nm | 
                                              (log2FoldChange > 4.5 & padj < 0.05)), 
                        aes(label = gene), 
                        min.segment.length = 0, max.overlaps = Inf) + 
        theme_cowplot() + 
        labs(x = "Mean expression", y = "log2(fold-change)", 
             title = paste0(nm, "\U141E")) + 
        theme(legend.position = "none",
              plot.title = element_text(face = "italic"))
})

## Correlation between log2 fold changes for ntu1/ntu2
stopifnot(all(rownames(resList$ntu1) == rownames(resList$ntu2)))
dfcorr <- data.frame(gene = rownames(resList$ntu1), 
                     ntu1 = resList$ntu1$log2FoldChange,
                     ntu2 = resList$ntu2$log2FoldChange)
corrplot <- list(
    ggplot(dfcorr, aes(x = ntu1, y = ntu2)) + 
        geom_abline(slope = 1, intercept = 0, linetype = "dashed", 
                    color = "grey50") + 
        geom_point(alpha = 0.25) + 
        geom_text_repel(data = dfcorr |>
                            dplyr::filter(gene %in% c("tna1", "SPCC576.17c")), 
                        aes(label = gene), 
                        min.segment.length = 0, max.overlaps = Inf) + 
        labs(x = "*ntu1\U141E* log2(fold-change)",
             y = "*ntu2\U141E* log2(fold-change)") + 
        theme_cowplot() + 
        theme(axis.title = element_markdown())
)

cowplot::plot_grid(
    plotlist = maplots, nrow = 1, align = "hv", axis = "b"
)
```

```{r rnaseq-genomeplot}

## Add information about gene location to result lists
genepos <- rowRanges(dds)
mcols(genepos) <- NULL
genepos <- as.data.frame(genepos)
resList <- lapply(resList, function(rl) {
    stopifnot(rownames(rl) == rownames(genepos))
    cbind(rl, genepos) |>
        as.data.frame() |>
        tibble::rownames_to_column("gene")
})

## Subset to region of interest
chrom <- genepos["tna1", "seqnames"]
startpos <- genepos["tna1", "start"]
endpos <- genepos["tna1", "end"]
windowsize <- 30000

resListToPlot <- lapply(resList, function(rl) {
    rl |> dplyr::filter(seqnames == chrom & 
                            start > startpos - windowsize & 
                            end < endpos + windowsize)
})
plotList <- lapply(names(resListToPlot), function(nm) {
    ggplot(resListToPlot[[nm]], 
           aes(x = start, y = log2FoldChange, 
               color = abs(log2FoldChange) > 3 & padj < 0.001)) + 
        geom_rect(aes(xmin = endpos, xmax = startpos, ymin = -Inf, ymax = Inf),
                  color = lighten(main_colors[4], amount = 0.6), fill = NA,
                  alpha = 0.1, show.legend = FALSE) + 
        geom_point(alpha = 0.6, size = 2) + 
        scale_color_manual(values = c(`FALSE` = "grey50", 
                                      `TRUE` = main_colors[4])) + 
        geom_text_repel(aes(label = ifelse(abs(log2FoldChange) > 3 & padj < 0.001,
                                           as.character(gene), "")),
                        size = 2.5, max.overlaps = 30, force = 10) + 
        scale_x_continuous(labels = scales::label_number(scale = 1e-6)) + 
        scale_y_continuous(limits = c(-7, 7)) + 
        labs(x = "Genome position (Mb)", y = "log2(fold-change)",
             title = paste0(nm, "\U141E")) + 
        theme_cowplot() + 
        theme(legend.position = "none",
              plot.title = element_text(face = "italic"))
})
genomeplot <- cowplot::plot_grid(plotlist = plotList, ncol = 1)
```

## Coverage tracks

```{r}
#| fig.width: 8
#| fig.height: 8
#| message: false
#| warning: false

covs_rna <- read.csv(params$rnatracks_ko) |>
    left_join(data.frame(sample = colnames(rnaseq_ko),
                         libsize = colSums(assay(rnaseq_ko, "counts")))) |>
    mutate(cpm = score / libsize * 1e6)
covs_chip <- read.csv(params$chiptracks_ko)
gtf <- rtracklayer::import(params$gtf)

goi <- c("med6", "SPAC1002.16c", "urg2", "urg3", "SPCC576.17c", "SPNCRNA.1263")
gtf <- subset(gtf, (gene_id %in% goi | gene_name %in% goi) & type == "gene") |>
    as.data.frame() |>
    left_join(idmap, by = join_by("gene_id" == "gene_stable_id")) |>
    dplyr::select(seqnames, start, end, strand, unique_einprot_id) |>
    mutate(y = ifelse(unique_einprot_id == "SPCC576.17c", 0, 1)) |>
    mutate(start2 = ifelse(strand == "+", start, end),
           end2 = ifelse(strand == "+", end, start)) |>
    mutate(unique_einprot_id_label = ifelse(grepl("^SP", unique_einprot_id), 
                                            unique_einprot_id, 
                                            paste0("*", unique_einprot_id, "\U207A*")))

## Create data frames for plotting
plotdf_chip1 <- covs_chip |>
    mutate(sample = sub("_rep[0-9]", "", sample)) |>
    mutate(exptype = stringr::str_extract(sample, "ChIP|IN")) |>
    mutate(pair = sub("_ChIP|_IN", "", sample)) |>
    mutate(flag = .capitalize(sub("_*TAG", "", sub(".*_KO_", "", pair)))) |>
    mutate(label = ifelse(pair == "ntu1_KO_ntu2TAG", "*ntu1\U141E*",
                          ifelse(pair == "ntu1_TAG", "*ntu2\U207A*",
                                 ifelse(pair == "ntu2_TAG", "*ntu1\U207A*",
                                        ifelse(pair == "ntu2_KO_ntu1TAG", "*ntu2\U141E*", 
                                               ""))))) |>
    mutate(cpm = score) |>
    mutate(pos = start) |>
    group_by(flag, seqnames) |>
    mutate(cpmnorm = cpm / max(abs(cpm))) |>
    ungroup() |>
    filter(pair %in% c("ntu2_TAG", "ntu2_KO_ntu1TAG", 
                       "ntu1_TAG", "ntu1_KO_ntu2TAG")) |>
    mutate(pair = factor(pair, levels = c("ntu1_TAG", "ntu2_KO_ntu1TAG",
                                          "ntu2_TAG", "ntu1_KO_ntu2TAG")))
plotdf_rna <- covs_rna |>
    filter(sample %in% c("SPBC530.08_KO_mRNA_rep1",
                         "SPBC16G5.16_KO_mRNA_rep1",
                         "wt_KO_mRNA_rep1")) |>
    mutate(sample = sub("_KO_mRNA_rep[0-9]", "", sample)) |>
    mutate(sample = replace(sample, sample == "SPBC16G5.16", "ntu1\U141E")) |>
    mutate(sample = replace(sample, sample == "SPBC530.08", "ntu2\U141E")) |>
    mutate(gene = sample) |>
    group_by(seqnames) |>
    mutate(cpmnorm = cpm / max(abs(cpm))) |>
    ungroup() |>
    mutate(sample = factor(sample, 
                           levels = c("wt", "ntu1\U141E", "ntu2\U141E")))
plotdf_label <- covs_rna |>
    filter(sample == "wt_KO_mRNA_rep1") |>
    mutate(cpmnorm = 0)

.get_empty_label <- function(x) {
    c("ntu1_TAG" = "", "ntu2_KO_ntu1TAG" = "",
      "ntu2_TAG" = "", "ntu1_KO_ntu2TAG" = "")
}
.add_space_to_label <- function(x) {
    paste(x, "    ")
}

maxIII <- max(plotdf_label$pos[plotdf_label$seqnames == "III"])
g1c <- ggplot(plotdf_label,
              aes(x = pos, y = cpmnorm)) + 
    geom_line(color = "white") + 
    geom_segment(x = maxIII - 2000, xend = maxIII,
                 y = 0, yend = 0, color = "black") + 
    geom_text(data = data.frame(seqnames = "III",
                                x = maxIII - 1000,
                                y = 0,
                                label = "2kb"),
              aes(x = x, y = y, label = label),
              vjust = -1) + 
    facet_grid(~ seqnames, scales = "free_x", space = "free_x")
g1a <- ggplot(plotdf_chip1 |>
                  mutate(seqnames = paste0("chr", seqnames)), 
              aes(x = pos, y = cpmnorm)) + 
    geom_area(aes(group = exptype, fill = interaction(exptype, flag))) + 
    geom_richtext(data = plotdf_chip1 |> 
                      mutate(seqnames = paste0("chr", seqnames)) |>
                      select(flag, pair, seqnames, pos, label) |>
                      filter(seqnames == "chrI") |>
                      filter(pos == min(pos)),
                  aes(x = pos, y = 0.65, label = label),
                  fill = NA, label.color = NA,
                  label.padding = grid::unit(rep(0, 4), "pt"),
                  hjust = 0, size = 4) + 
    scale_fill_manual(values = c(`ChIP.Ntu1` = main_colors[5], 
                                 `IN.Ntu1` = na_color,
                                 `ChIP.Ntu2` = main_colors[1], 
                                 `IN.Ntu2` = na_color)) + 
    facet_nested(flag + pair ~ seqnames, scale = "free_x", space = "free_x",
                 switch = "y",
                 labeller = labeller(pair = .get_empty_label, 
                                     flag = .add_space_to_label),
                 nest_line = element_line(),
                 strip = strip_nested(size = "variable")) + 
    scale_y_continuous(limits = c(0, 0.9), expand = c(0, 0))
g1b <- ggplot(plotdf_rna, 
              aes(x = pos, y = cpmnorm)) + 
    geom_area(aes(group = strand, fill = gene)) + 
    scale_fill_manual(values = c(wt = lighten("black", amount = 0.4), 
                                 "ntu1\U141E" = lighten(main_colors[5], amount = 0.4), 
                                 "ntu2\U141E" = lighten(main_colors[1], amount = 0.4))) + 
    facet_grid(sample ~ seqnames, scale = "free_x", space = "free_x",
               switch = "y") + 
    scale_y_continuous(limits = c(-0.9, 0.3), expand = c(0, 0))
g2 <- ggplot(gtf, aes(x = start2, xend = end2, y = y, yend = y, 
                      color = unique_einprot_id %in% c("tna1", "SPCC576.17c"), 
                      label = unique_einprot_id_label)) + 
    geom_segment(arrow = grid::arrow(length = unit(0.05, "inches")), 
                 linewidth = 2) + 
    geom_richtext(
        aes(x = (start2 + end2) / 2, y = y - 0.5),
        fill = NA, label.color = NA,
        label.padding = grid::unit(rep(0, 4), "pt"),
        size = 4
    ) + 
    # geom_text(aes(x = (start2 + end2) / 2, y = y - 0.5), 
    #           size.unit = "in", size = 0.1) + 
    scale_color_manual(values = c(`FALSE` = "black", `TRUE` = main_colors[4])) + 
    facet_grid(~ seqnames, scales = "free_x", space = "free_x") + 
    ylim(-0.75, 1.5)
gcov <- cowplot::plot_grid(
    g1c + theme_void() + theme(strip.text = element_blank()),
    g1a + theme_void() + theme(legend.position = "none", 
                               strip.text.y = element_text(size = 12),
                               strip.text.x = element_text(size = 12, hjust = 0.95)), 
    NULL,
    g1b + theme_void() + theme(strip.text.y = element_text(size = 12, 
                                                           face = "italic"), 
                               legend.position = "none",
                               strip.text.x = element_blank()), 
    g2 + theme_void() + 
        theme(legend.position = "none",
              strip.text = element_blank()), 
    ncol = 1, rel_heights = c(0.75, 2, 0.5, 2, 1),
    align = "v", axis = "lr")
(gcov2 <- cowplot::plot_grid(
    g1c + theme_bw() + theme(axis.text.x = element_text(size = 5)),
    g1a + theme_bw() + theme(legend.position = "none",
                             axis.text.x = element_text(size = 5)), 
    g1b + theme_bw() + theme(legend.position = "none",
                             axis.text.x = element_text(size = 5)), 
    g2 + theme_bw() + 
        theme(legend.position = "none",
              axis.text.x = element_text(size = 5)), 
    ncol = 1, rel_heights = c(0.75, 2, 2, 1),
    align = "v", axis = "lr"))
```

## Put together

```{r Fig7}
#| fig.width: 13
#| fig.height: 14
#| warning: false

fig7b <- ggdraw() + 
    draw_image("schematics/Fig7B_Nattou_protein_domains_cartoon_Arial.png")
fig7e <- ggdraw() + 
    draw_image("schematics/Fig7E_Nattou_lacR_cartoon_images_Arial.png")
fig7f <- ggdraw() + 
    draw_image("schematics/Fig7F_Nattou_lacR_microscopy_images_Arial.png")

fig7c <- wrap_plots(maplots, ncol = 1) | corrplot[[1]]

cowplot::plot_grid(
    cowplot::plot_grid(
        gg_ipdimer, 
        fig7c,
        nrow = 1, 
        rel_widths = c(1, 2), 
        labels = c("A", "C"),
        align = "h", axis = "bt"
    ),
    cowplot::plot_grid(
        fig7b, 
        gcov,
        rel_widths = c(1, 1),
        labels = c("B", "D")
    ),
    cowplot::plot_grid(
        fig7e,
        NULL,
        fig7f,
        NULL,
        p_zone + theme(legend.position = "top"), 
        labels = c("E", "", "F", "", "G"),
        # rel_widths = c(7.65, 0.5, 1.45, 0.5, 9)
        rel_widths = c(6, 0.75, 1.2, 0.75, 10.4), nrow = 1
    ), 
    ncol = 1
)
```

## Supplementary Figure

```{r suppFig7-chip}
#| warning: false
#| message: false

plotdf_chip2 <- covs_chip |>
    mutate(sample = sub("_rep[0-9]", "", sample)) |>
    mutate(exptype = stringr::str_extract(sample, "ChIP|IN")) |>
    mutate(pair = sub("_ChIP|_IN", "", sample)) |>
    mutate(flag = sub("_*TAG", "", sub(".*_KO_", "", pair))) |>
    mutate(label = ifelse(pair == "ntu2_delTMD", "delTMD",
                          ifelse(pair == "ntu2_TAG", "ntu2+", ""))) |>
    mutate(cpm = score) |>
    mutate(pos = start) |>
    group_by(seqnames) |>
    mutate(cpmnorm = cpm / max(abs(cpm))) |>
    ungroup() |>
    filter(pair %in% c("ntu2_TAG", "ntu2_delTMD")) |>
    mutate(pair = ifelse(pair == "ntu2_delTMD", "Ntu2\U141E\U1D40\U1D39\U1D30", 
                         ifelse(pair == "ntu2_TAG", "Ntu2", pair))) |>
    mutate(pair = factor(pair, levels = c("Ntu2", "Ntu2\U141E\U1D40\U1D39\U1D30")))

gtr <- ggplot(plotdf_chip2 |>
                  mutate(seqnames = paste0("chr", seqnames)), 
              aes(x = pos, y = cpmnorm)) + 
    geom_area(aes(group = exptype, fill = interaction(exptype, flag))) + 
    scale_fill_manual(values = c(ChIP.ntu2_delTMD = main_colors[3], 
                                 IN.ntu2_delTMD = na_color,
                                 ChIP.ntu2 = main_colors[1], 
                                 IN.ntu2 = na_color)) + 
    facet_grid(pair ~ seqnames, scale = "free_x",
               space = "free_x", switch = "y") + 
    scale_y_continuous(limits = c(0, 0.9), expand = c(0, 0))
gcovdel <- cowplot::plot_grid(
    gtr + theme_void() + theme(strip.text.y = element_text(size = 12, vjust = 0.05),
                               legend.position = "none", 
                               strip.text.x = element_text(size = 12, hjust = 0.95)), 
    g2 + theme_void() + 
        theme(legend.position = "none",
              strip.text = element_blank()), 
    ncol = 1, rel_heights = c(2, 1),
    align = "v", axis = "lr")
```

```{r suppFig7}
#| fig.width: 13
#| fig.height: 15
#| warning: false

figs6a <- ggdraw() + 
    draw_image("schematics/FigS7A_Nattou_AF2_fl_Arial.png")

cowplot::plot_grid(
    figs6a,
    cowplot::plot_grid(
        p_rel_dis,
        genomeplot_tpm,
        nrow = 1, 
        rel_widths = c(1, 1), 
        labels = c("B", "C")
    ),
    cowplot::plot_grid(
        genomeplot,
        gcovdel,
        # maplots$ID83delTMD,
        nrow = 1, 
        rel_widths = c(1, 1),
        labels = c("C", "E")
    ),
    ncol = 1, rel_heights = c(1, 1, 1)
)
```

# Session info

<details>
<summary><b>
Session info
</b></summary>
```{r}
sessionInfo()
```
</details>