### 2024-07-09 index.qmd
Add some ORCIDs, reduce author/affiliation font size, update figure links

---
title: "A comprehensive Schizosaccharomyces pombe atlas of physical transcription factor interactions with proteins and DNA"
subtitle: "Analysis code for figures"
date: today
author:
  - name: Merle Skribbe
    email: merle.skribbe@gmail.com
    affiliations:
      - ref: fmi
      - ref: unibas
  - name: Charlotte Soneson
    orcid: 0000-0003-3833-2169
    email: charlotte.soneson@fmi.ch
    affiliations:
      - ref: fmi
      - ref: sib
  - name: Michael B Stadler
    orcid: 0000-0002-2269-4934
    email: michael.stadler@fmi.ch
    affiliations:
      - ref: fmi
      - ref: unibas
      - ref: sib
  - name: Michaela Schwaiger
    affiliations:
      - ref: fmi
      - ref: sib
  - name: Vishnu Narayanan Suma Sreechakram
    affiliations:
      - ref: giessen
  - name: Vytautas Iesmantavicius
    orcid: 0000-0002-2512-9957
    affiliations:
      - ref: fmi
  - name: Daniel Hess
    orcid: 0000-0002-1642-5404
    affiliations:
      - ref: fmi
  - name: Eliza Pandini Figueiredo Moreno
    affiliations:
      - ref: fmi
  - name: Sigurd Braun
    affiliations:
      - ref: giessen
  - name: Jan Seebacher
    orcid: 0000-0002-7858-2720
    affiliations:
      - ref: fmi
  - name: Sebastien Smallwood
    affiliations:
      - ref: fmi
  - name: Marc Bühler
    orcid: 0000-0001-6661-9795
    email: marc.buehler@fmi.ch
    affiliations:
      - ref: fmi
      - ref: unibas
affiliations:
  - id: fmi
    name: Friedrich Miescher Institute for Biomedical Research
    address: Fabrikstrasse 24
    postal-code: 4056
    city: Basel
    country: Switzerland
  - id: unibas
    name: University of Basel
    address: Petersplatz 10
    postal-code: 4003
    city: Basel
    country: Switzerland
  - id: sib
    name: SIB Swiss Institute of Bioinformatics
    address: Fabrikstrasse 24
    postal-code: 4056
    city: Basel
    country: Switzerland
  - id: giessen
    name: Institute for Genetics, Justus-Liebig-University Giessen
    city: Giessen
    country: Germany
license: MIT License
format: 
    html:
        embed-resources: true
        fig-cap-location: top
---

<style>
.author {
  font-size: 0.7em;
}

.affiliation {
  font-size: 0.7em;
}
</style>

## Interactive data exploration with `TFexplorer`

## Links to raw data

## Links to publication

## Links to code for generating figures

All source code is available from GitHub at [https://github.com/fmicompbio/Spombe_TFome](https://github.com/fmicompbio/Spombe_TFome).
Click on a specific figure below to jump directly to the corresponding sources.

|          |                                                                                                  |     |
|----------|----------------------------|----------------------------|
| Figure 1 | [![](ms_figure_1_overview_files/figure-html/Fig1-1.png){width="4cm"}](ms_figure_1_overview.html) |  [![](ms_figure_1_overview_files/figure-html/suppFig1-1.png){width="4cm"}](ms_figure_1_overview.html#Supplementary_figure)   |
| Figure 2 | [![](ms_figure_2_overview-heatmaps_files/figure-html/Fig2-1.png){width="4cm"}](ms_figure_2_overview-heatmaps.html) | [![](ms_figure_2_overview-heatmaps_files/figure-html/suppFig2-1.png){width="4cm"}](ms_figure_2_overview-heatmaps.html#Supplementary_figure)    |
| Figure 3 | [![](ms_figure_3_chip_files/figure-html/Fig3-1.png){width="4cm"}](ms_figure_3_chip.html) | [![](ms_figure_3_chip_files/figure-html/suppFig3-1.png){width="4cm"}](ms_figure_3_chip.html#Supplementary_figure)    |
| Figure 4 | [![](ms_figure_4_motifs_files/figure-html/Fig4-1.png){width="4cm"}](ms_figure_4_motifs.html) | [![](ms_figure_4_motifs_files/figure-html/suppFig4-1.png){width="4cm"}](ms_figure_4_motifs.html#Supplementary_figure)    |
| Figure 5 | [![](ms_figure_5_ipms_files/figure-html/Fig5-1.png){width="4cm"}](ms_figure_5_ipms.html) | [![](ms_figure_5_ipms_files/figure-html/suppFig5-1.png){width="4cm"}](ms_figure_5_ipms.html#Supplementary_figure)    |
| Figure 6 | [![](ms_figure_6_networks_files/figure-html/Fig6-1.png){width="4cm"}](ms_figure_6_networks.html) | [![](ms_figure_6_networks_files/figure-html/suppFig6-1.png){width="4cm"}](ms_figure_6_networks.html#Supplementary_figure)    |
| Figure 7 | [![](ms_figure_7_heterodimer_files/figure-html/Fig7-1.png){width="4cm"}](ms_figure_7_heterodimer.html) | [![](ms_figure_7_heterodimer_files/figure-html/suppFig7-1.png){width="4cm"}](ms_figure_7_heterodimer.html#Supplementary_figure)    |