### 2024-07-15 ms_figure_4_motifs.Rmd
rename Fig3 to Fig4

---
title: "Figure 4 - Motifs"
author: "Michael Stadler, Charlotte Soneson, Merle Skribbe"
date: "`r Sys.Date()`"
output: 
    html_document:
        theme: united
        toc: true
        toc_float: true
        code_folding: hide
params:
    dbdtxt: "data/TF_DBD_annotation.txt"
    motifrds: "data/motifs_annotated.rds"
    peakcsv: "data/fused_peaks_filtered.csv.gz"
    genomefasta: "reference/Schizosaccharomyces_pombe.ASM294v2.dna.toplevel.fa"
    ncpu: 4
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = c("png", "cairo_pdf"))
```

# Parameter values

```{r params}
params
```

# Load required packages

```{r load-packages}
suppressPackageStartupMessages({
    library(circlize)
    library(ComplexHeatmap)
    library(grid)
    library(ggplot2)
    library(cowplot)
    library(dplyr)
    library(tidyr)
    library(universalmotif)
    library(TFBSTools)
    library(GenomicRanges)
    library(parallel)
    library(colorspace)
    library(forcats)
    library(monaLisa)
    library(ggrepel)
    library(Biostrings)
    library(ggupset)
})

source("params/plot_settings.R")

# capitalize the first letter of a string
.capitalize <- function(x) {
    paste0(toupper(substr(x, 1, 1)), substr(x, 2, nchar(x)))
}
```

# Read data

We start by reading all tables and objects needed for this figure.
Factor levels are defined in the order they should appear in the figure
panels, and peak sequences are extracted from the genome.

Remark: Motifs are read from the `r params$motifrds` file, which is in
`.rds` format that can be used by `R`. The motifs are also available in the
file [`data/motifs.meme`](data/motifs.meme), which is a plain text file with the
motifs in the MEME format: https://meme-suite.org/meme/doc/meme-format.html.

```{r load-data}
# `dbd`: Transcription factor DNA binding domain information
dbd <- read.delim(params$dbdtxt) |>
    dplyr::mutate(cons = factor(ifelse(Conserved_human, "yes", "no"),
                                levels = c("yes", "no")),
                  DBD_class_simple = factor(
                      ifelse(DBD_class == "Zn(II)2Cys6", "Zn(II)2Cys6", "other"),
                      levels = c("other", "Zn(II)2Cys6")))

# `motifs`: table with annotated motif
motifs <- readRDS(params$motifrds)
motifs$altname <- motifs$name

# ... extract motifs as TFBSTools::PFMatrixList
umL <- universalmotif::to_list(motifs)
pfmL <- do.call(TFBSTools::PFMatrixList,
                lapply(seq.int(nrow(motifs)), function(i) {
                    universalmotif::convert_motifs(motifs = umL[[i]],
                                                   class = "TFBSTools-PFMatrix")
                }))

# `gnm`: S. pombe reference genome sequence
gnm <- readDNAStringSet(params$genomefasta)
names(gnm) <- sub(" .*$", "", names(gnm))

# `peakgr`: ChIP-seq peaks as GRanges object
peakgr <- as(read.csv(params$peakcsv, row.names = 1), "GRanges")
peakgr$peaktype <- factor(.capitalize(peakgr$peaktype),
                          levels = c("Common peaks (ubiquitous)",
                                     "Common peaks (frequent)",
                                     "Specific peaks"))
peakgr_specific <- peakgr[peakgr$peaktype == "Specific peaks"]

# ... extract peak enrichment matrix
is_enr <- as.matrix(mcols(peakgr)[, grep("^is_enr_in[.]", colnames(mcols(peakgr)))])
colnames(is_enr) <- sub("^is_enr_in.", "", colnames(is_enr))
is_enr_specific <- as.matrix(mcols(peakgr_specific)[, grep("^is_enr_in[.]", colnames(mcols(peakgr_specific)))])
colnames(is_enr_specific) <- sub("^is_enr_in.", "", colnames(is_enr_specific))

# ... extract peak sequences from genome
peakseqs <- getSeq(gnm, peakgr)
peakseqs_specific <- getSeq(gnm, peakgr_specific)
```

# Helper functions

Here we define some helper functions that are used to generate this figure:

- `calcKmerEnr`: Counts the frequencies of k-mer words of length `k` in
  a set of sequences (`seqs`) that are grouped into foreground/background sets
  by `fg`, and calculates enrichment values for each k-mer word comparing
  the foreground frequencies to either the background sequences
  (`bg_type = "other"`) or to a Markov model of order `Markov_order` that is
  estimated from the foreground sequences.
- `kmerEnrPlot`: Calls `calcKmerEnr` to obtain k-mer enrichments and visualizes
  them in a scatter plot (each point corresponding to a k-mer), with x- and y-axis
  coordinates corresponding to the log k-mer frequency in foreground sequences
  and the log k-mer enrichment, respectively.
- `create_cobound_peaks_plot`: Upset plot showing co-bound peaks for a set of TFs.

```{r helpers}
calcKmerEnr <- function(k = 5, # k-mer length
                        seqs,  # DNAStringSet
                        fg,    # logical vector of foreground in `seqs`
                        bg_type = c("other", "Markov"),
                        Markov_order = 1,
                        include_revcomp = TRUE,
                        ... # forwarded to monaLisa::getKmerFreq
                        ) {
    # digest arguments
    bg_type <- match.arg(bg_type)
    stopifnot(exprs = {
        is.numeric(k) && length(k) == 1L
        is(seqs, "DNAStringSet")
        is.logical(fg) && length(fg) == length(seqs)
        is.numeric(Markov_order) && length(Markov_order) == 1L
    })

    # fg
    fg_kmers <- monaLisa::getKmerFreq(seqs[fg], kmerLen = k,
                                      MMorder = Markov_order,
                                      includeRevComp = include_revcomp,
                                      ...)
    fg_freq <- fg_kmers$freq.obs
    kmers <- names(fg_freq)

    # bg
    if (identical(bg_type, "other")) {
        bg_kmers <- monaLisa::getKmerFreq(seqs[!fg], kmerLen = k,
                                          MMorder = Markov_order,
                                          includeRevComp = include_revcomp,
                                          ...)
        bg_freq <- bg_kmers$freq.obs
    } else if (bg_type == "Markov") {
        bg_freq <- fg_kmers$freq.exp
    }

    # norm
    N <- min(sum(fg_freq), sum(bg_freq))
    fg_freq_norm <- fg_freq / sum(fg_freq) * N
    bg_freq_norm <- bg_freq / sum(bg_freq) * N

    # return results
    return(list(kmers = names(fg_freq),
                fg_kmers = fg_kmers,
                fg_freq = fg_freq,
                fg_freq_norm = fg_freq_norm,
                bg_kmers = bg_kmers,
                bg_freq = bg_freq,
                bg_freq_norm = bg_freq_norm,
                pearsonresid = (fg_freq_norm - bg_freq_norm) /
                         sqrt(bg_freq_norm),
                nCG = vcountPattern("C", kmers) + vcountPattern("G", kmers)))
}

kmerEnrPlot <- function(k = 5, # k-mer length
                        seqs,  # DNAStringSet
                        fg,    # logical vector of foreground in `seqs`
                        bg_type = c("other", "Markov"),
                        Markov_order = 1,
                        colorByMotif = NULL,
                        nlabel = 0,
                        title = NULL,
                        auto_subtitle = FALSE,
                        jitter_x = FALSE,
                        include_revcomp = TRUE,
                        font_size = 14,
                        ... # forwarded to monaLisa::getKmerFreq
                        ) {
    # digest arguments
    stopifnot(exprs = {
        is.null(colorByMotif) || is(colorByMotif, "PFMatrixList")
        is.numeric(nlabel) && length(nlabel) == 1L
        is.null(title) || (is.character(title) && length(title) == 1L)
    })

    # calculate k-mer enrichments
    enr <- calcKmerEnr(k = k,
                       seqs = seqs,
                       fg = fg,
                       bg_type = bg_type,
                       Markov_order = Markov_order,
                       include_revcomp = include_revcomp)

    # enrichment plot
    jitter_amount <- if (jitter_x) 0.5 else 0
    pd <- data.frame(kmer = enr$kmers,
                     fg_freq = enr$fg_freq,
                     bg_freq = enr$bg_freq,
                     fg_freq_norm = enr$fg_freq_norm,
                     bg_freq_norm = enr$bg_freq_norm,
                     pearsonresid = enr$pearsonresid,
                     nCG = enr$nCG,
                     jitter_x = runif(length(enr$kmers),
                                      min = -jitter_amount, max = jitter_amount))
    if (!is.null(colorByMotif)) {
        pd$score <- monaLisa::motifKmerSimilarity(x = colorByMotif, kmerLen = k,
                                                  includeRevComp = include_revcomp)[1, ]
    }

    # ... base plot
    p <- ggplot(pd, aes(fg_freq + 1 + jitter_x, pearsonresid)) +
        scale_x_log10() +
        labs(title = if (is.null(title)) element_blank() else title,
             x = paste0(k, "-mer frequency in peaks + 1"),
             y = paste0(k, "-mer enrichment")) +
        theme_cowplot(font_size) +
        theme(legend.position = "none")
    # ... subtitle
    if (auto_subtitle) {
        if (identical(bg_type, "Markov")) {
            p <- p + labs(subtitle = paste0(k, "-mer enrichments (background: ",
                                            Markov_order,
                                            "th-order Markov model)"))
        } else if (identical(bg_type, "other")) {
            p <- p + labs(subtitle = paste0(k, "-mer enrichments (background: other peaks)"))
        }
    }
    # ... points
    if (is.null(colorByMotif)) {
        p <- p + geom_point(color = "gray62", alpha = 0.5, size = 1.5 * font_size / 14) +
            geom_hline(yintercept = 0, linetype = "dashed")
    } else {
        p <- p +
            geom_point(mapping = aes(color = score^0.4), alpha = 0.5) +
            scale_colour_gradientn(colors = hcl.colors(11, "Oslo")[-1]) +
            geom_hline(yintercept = 0, linetype = "dashed")
    }
    # ... point labels
    if (nlabel > 0) {
        p <- p + geom_text_repel(
            # data = pd[order(pd$pearsonresid,
            #                 decreasing = TRUE)[seq.int(nlabel)], ],
            data = pd |>
                arrange(desc(pearsonresid)) |>
                mutate(label_i = seq_along(pearsonresid),
                       kmerrc = as.character(reverseComplement(DNAStringSet(kmer))),
                       labelrc_i = match(kmer, kmerrc),
                       label_str = ifelse(label_i <= nlabel & label_i <= labelrc_i, kmer, "")),
            mapping = aes(label = label_str), min.segment.length = 0,
            size = font_size / ggplot2::.pt, color = "black",
            max.overlaps = Inf)
    }

    # return plot
    return(p)
}

create_cobound_peaks_plot <- function(sel_tfs) {
    stopifnot(all(sel_tfs %in% colnames(is_enr)))

    ovL <- as.list(as.data.frame(is_enr[, sel_tfs]))
    ovL <- lapply(ovL, function(i) rownames(is_enr)[i])

    pd <- stack(ovL) |>
        group_by(values) |>
        summarise(ind = list(ind)) |>
        mutate(peaktype = peakgr[values]$peaktype)

    ggplot(pd, aes(x = ind)) +
        geom_bar(aes(fill = peaktype)) +
        scale_fill_manual(values = peakset_colors) +
        scale_x_upset() +
        labs(x = element_blank(), y = "Number of peaks", fill = "Peak type:  ") +
        theme_cowplot(12) +
        theme(axis.ticks.x = element_blank(),
              legend.position = "bottom") +
        theme_combmatrix(combmatrix.label.extra_spacing = 5,
                         combmatrix.label.make_space = TRUE,
                         combmatrix.label.width = unit(18, "mm")) +
        guides(fill = guide_legend(nrow = 3))
}
```

# Figure 4

## Zn(II)2Cys6 family motifs

### Zn(II)2Cys6 TF motifs k-mer enrichment plots

Frequency and enrichments of 6-mer sequence words in ChIP-seq peaks from
transcription factors of the Zn(II)2Cys6 (GAL4) family:

```{r plot-ZnII2Cys6-kmer}
#| fig-width: 6
#| fig-height: 6
# select motifs of TFs from the Zn(II)2Cys6 family
sel_motifs <- motifs[motifs$family == "Zn(II)2Cys6", "name"]
tf_names <- sub("[.]m[0-9]+$", "", sel_motifs)

# select a single TF and motif of interest ("Toe3.m1")
sel_motif_idx_1 <- 7
motif_name1 <- sel_motifs[sel_motif_idx_1]

# plot 6-mer enrichments in enriched peaks from which the motif was identified
ZnII2Cys6_kmer_1 <- kmerEnrPlot(k = 6,
                                seqs = peakseqs_specific,
                                fg = is_enr_specific[, tf_names[sel_motif_idx_1]],
                                bg_type = "other",
                                title = sel_motifs[sel_motif_idx_1],
                                nlabel = 9,
                                colorByMotif = pfmL[match(sel_motifs[sel_motif_idx_1], name(pfmL))],
                                font_size = 10)

plot_grid(ZnII2Cys6_kmer_1)
```

### Zn(II)2Cys6 TF `CCG`, `CGG` enrichments (TF peaks from `r length(sel_motifs)` selected motifs)

```{r hm-ZnII2Cys6-selected-3mer-enr}
#| fig-width: 8
#| fig-height: 4

# select Zn(II)2Cys6 TFs
tf_names <- c("Cha4", "Mca1", "Moc3", "Pho7", "Prt1", 
              "Grt1", "Gsf1", "Thi1", "Thi5", "Toe1", 
              "Toe2", "Toe3", "Toe4", "SPAC11D3.11c", "SPAC1327.01c", 
              "SPAC1F7.11c", "SPAC25B8.11", "SPAC2H10.01", "SPAC3C7.04", 
              "SPAC3H8.08c", "SPBC1348.12", "Ntu1", "SPBC16G5.17", 
              "SPBC1773.12", "SPBC1773.16c", "Ntu2", "SPBC56F2.05c", 
              "SPCC320.03", "SPCC417.09c", "SPCC757.04", "SPCC777.02", 
              "SPCC965.10") 
length(tf_names)

# filter out not ChIP'ed TFs or TFs with less than `min_peaks_specific` specific peaks
min_peaks_specific <- 3
n_peaks_specific <- colSums(
    is_enr_specific[, intersect(colnames(is_enr_specific), tf_names)]
    )[tf_names]
tf_names <- tf_names[!is.na(n_peaks_specific) & n_peaks_specific >= min_peaks_specific]
length(tf_names)

# calculate 3-mer enrichments
enr_list_3mers <- list()
for (i in seq_along(tf_names)) {
    enr_list_3mers[[i]] <- calcKmerEnr(k = 3,
                                       seqs = peakseqs_specific,
                                       fg = is_enr_specific[, tf_names[i]],
                                       bg_type = "other")
}

enr_3mers <- do.call(rbind, lapply(enr_list_3mers, "[[", "pearsonresid"))
rownames(enr_3mers) <- tf_names

# plot heatmap
# ... define heatmap font sizes
fs <- 5.5 # small
fl <- 8   # large
ft <- 10  # title

# ... calculate data range for mapping to colors
mx <- max(abs(enr_3mers))
qs <- quantile(abs(enr_3mers), .99)

# ... prepare heatmap annotations:
#      nCG: number of G+C bases in the 3-mer word
#      highlight: emphasize CCG, CGG, GGC and GCC
kmer_annot <- columnAnnotation(
    nCG = vcountPattern("S", DNAStringSet(colnames(enr_3mers)), fixed = FALSE),
    highlight = ifelse(colnames(enr_3mers) %in% c("CCG", "CGG", "GGC", "GCC"), "1", "0"),
    col = list(highlight = c("0" = bg_color, "1" = "black"),
               nCG = structure(hcl.colors(5, "Greens")[-5], names = 3:0)),
    show_legend = c(TRUE, FALSE), show_annotation_name = c(FALSE, FALSE),
    annotation_name_gp = gpar(fontsize = fl),
    annotation_legend_param = list(labels_gp = gpar(fontsize = fl),
                                   ncol = 4,
                                   title = "Number of G+C   ",
                                   title_position = "lefttop",
                                   title_gp = gpar(fontsize = fl)),
    simple_anno_size = unit(3, "mm"))

# ... generate main heatmap with annotations
hm_3mers <- Heatmap(enr_3mers, name = "Pearson's resid.",
                    col = circlize::colorRamp2(
                        breaks = c(-mx, seq(-qs, qs, length.out = 62), mx),
                        colors = colorRampPalette(enrichment_heatmap_colors)(64)),
                    cluster_rows = TRUE, show_row_dend = FALSE,
                    column_dend_reorder = match(colnames(enr_3mers),
                                                c("GCG", "CCG", "GGC", "GGG"),
                                                nomatch = 100),
                    cluster_columns = TRUE, column_dend_height = unit(10, "mm"),
                    row_names_gp = gpar(fontsize = fl, fontfamily = "sans"),
                    column_names_gp = gpar(fontsize = fs, fontfamily = "mono"),
                    column_title = "Binuclear zinc cluster TFs",
                    column_title_side = "top", column_title_gp = gpar(fontsize = ft),
                    border = TRUE, border_gp = gpar(lwd = 0.5),
                    show_row_names = TRUE, show_column_names = TRUE,
                    use_raster = FALSE, show_heatmap_legend = TRUE,
                    bottom_annotation = kmer_annot,
                    heatmap_legend_param = list(at = round(c(-mx, 0, mx), 1),
                                                labels_gp = gpar(fontsize = fl),
                                                border = "gray10",
                                                legend_direction = "horizontal",
                                                legend_width = unit(25, "mm"),
                                                title = "3-mer enrichment   ",
                                                title_position = "lefttop",
                                                title_gp = gpar(fontsize = fl)),
                    width = unit(130, "mm"), # heatmap body width
                    height = unit(3 * nrow(enr_3mers), "mm"))

# draw the heatmap and grab as a grid graphics `grob`
# (needed for combining figure panels)
# use defined random number seed to make clustering deterministic
set.seed(42L)
hm_ZnII2Cys6_sel_3mers <- grid.grabExpr(
    hm_3mers <- draw(hm_3mers, merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "right"),
    width = 40, height = 140
)

plot_grid(hm_ZnII2Cys6_sel_3mers)
```

```{r peak-overlaps-within-tf-family}
#| fig.width: 10
#| fig.height: 8

# get enrichment calls for specific peaks that are
# bound by at least one of the binuclear zinc cluster TFs used above
pd <- as.data.frame(is_enr_specific[, tf_names]) |>
    tibble::rownames_to_column("peak_name") |>
    pivot_longer(cols = all_of(tf_names),
                 values_to = "enriched",
                 names_to = "tf_name") |>
    filter(enriched) |>
    group_by(peak_name) |>
    summarise(tfs_enr = list(tf_name))

# create upset plot
upset_ZnII2Cys6_sel <- ggplot(pd, aes(x = tfs_enr)) +
        geom_bar() +
        scale_x_upset() +
        labs(x = element_blank(),
             y = "Number of enriched peaks") +
        theme_cowplot()
upset_ZnII2Cys6_sel
```

## Known and new motif characterization

Display information for selected motifs (`sel_motifs`, containing examples for 
both previously known and novel motifs):

```{r plot-known-new-details}
#| fig-width: 12
#| fig-height: 2.75
#| warning: false

# select motifs for detailed visualization
sel_motifs <- c( "Toe2.m1", "Esc1.m1", "SPAC3F10.12c.m1",
                 "Ams2.m1", "Zip1.m1", "Adn2.m4")
tf_names <- sub("[.]m[0-9]+$", "", sel_motifs)

# create motif visualizations
# ... setup plotting parameters
fgbg_col <- c(main_colors[1], main_colors[5]) # colors for foreground and background
gnm_col <- bg_color # color for NA or "other"
w <- 2500  # width in bases around peak-center for meta-profiles
fs <- 9 # font size
nlabel <- 1 # number of k-mer words to label in enrichment plot

# ... iterate over motifs
motif_characterization_plots <- list()
for (i in seq_along(sel_motifs)) {
    # get motif info
    m1 <- motifs$motif[match(sel_motifs[i], motifs$name)]
    is_fg_specific <- is_enr_specific[, tf_names[i]]
    fg_peak_names <- names(peakgr)[is_enr[, tf_names[i]]]
    bg_peak_names <- setdiff(names(peakgr), fg_peak_names)
    fgbg_num <- c(length(fg_peak_names), length(bg_peak_names))
    fgbg_names <- sprintf("%s (%d)", c("foreground", "background"), fgbg_num)
    fgbg_names_short <- sprintf("%d", fgbg_num)

    # scan genome sequence with motif
    hits_genome <- universalmotif::scan_sequences(motifs = m1,
                                                  sequences = gnm,
                                                  threshold = 1e-4,
                                                  threshold.type = "pvalue", 
                                                  RC = TRUE,
                                                  verbose = if (interactive()) 3 else 0,
                                                  nthreads = params$ncpu,
                                                  return.granges = TRUE)

    # seqlogo
    gg1_seqlogo <- universalmotif::view_motifs(
        motifs = m1,
        show.positions = FALSE) +
        labs(title = sel_motifs[i],
             subtitle = paste0(sum(is_fg_specific), " specific peaks")) +
        theme_cowplot(fs) +
        theme(axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              axis.line.x = element_blank(),
              legend.position = "none")

    # k-mer enrichments
    gg1_kmers <- kmerEnrPlot(k = 6,
                             seqs = peakseqs_specific,
                             fg = is_enr_specific[, tf_names[i]],
                             bg_type = "other",
                             nlabel = nlabel,
                             colorByMotif = pfmL[match(sel_motifs[i], name(pfmL))],
                             font_size = fs) +
        labs(x = "6-mer freq. + 1",
             y = "6-mer enr.") +
        theme(legend.position = "none")

    # motif location relative to peak mid
    nrst <- distanceToNearest(x = resize(hits_genome, width = 1L, fix = "center"),
                              subject = resize(peakgr, width = 1L, fix = "center"))
    nrst_fg <- nrst[subjectHits(nrst) %in% match(fg_peak_names, names(peakgr))]
    nrst_bg <- nrst[subjectHits(nrst) %in% match(bg_peak_names, names(peakgr))]
    relpos_fg <- mid(ranges(hits_genome))[queryHits(nrst_fg)] - mid(ranges(peakgr))[subjectHits(nrst_fg)]
    relpos_bg <- mid(ranges(hits_genome))[queryHits(nrst_bg)] - mid(ranges(peakgr))[subjectHits(nrst_bg)]
    pd_loc <- data.frame(motif = rep(sel_motifs[i], length(nrst_fg) + length(nrst_bg)),
                         relpos = c(relpos_fg, relpos_bg),
                         type = factor(rep(c(fgbg_names[1], fgbg_names[2]),
                                           c(length(nrst_fg), length(nrst_bg))),
                                       levels = fgbg_names))
    gg1_loc <- ggplot(pd_loc |> filter(abs(relpos) < w + 500),
                 aes(relpos / 1000, color = type)) +
        geom_density(linewidth = 1) +
        geom_vline(xintercept = 0, linetype = "dotted") +
        scale_y_continuous(n.breaks = 3) +
        scale_x_continuous(limits = c(-w, w) / 1000, expand = expansion(mult = 0)) +
        scale_color_manual(values = structure(fgbg_col, names = fgbg_names)) +
        annotate("text", x = w / 1000, y = Inf, hjust = 1, vjust = 1,
                 label = paste0(c("", "\n"), fgbg_names_short),
                 colour = fgbg_col, fontface = "bold", size = fs / ggplot2::.pt) +
        labs(x = "Position relative to peak center (kb)",
             y = "Density of motifs") +
        theme_cowplot(fs) +
        theme(legend.position = "none",
              legend.justification = c(1, 1))

    # motif hit enrichment in genomic regions
    ov <- findOverlaps(query = hits_genome, subject = peakgr)
    ov_type <- rep("non-peak", length(hits_genome))
    ov_type[queryHits(ov)] <- 
        ifelse(subjectHits(ov) %in% match(fg_peak_names, names(peakgr)),
               fgbg_names[1], fgbg_names[2])
    ov_type <- factor(ov_type,
                      levels = c("non-peak", fgbg_names[2], fgbg_names[1]))
    n_ov_obs <- unclass(table(hits_genome$motif, ov_type))
    n_ov_exp <- unclass(table(hits_genome$motif)) %*%
        t(structure(c(sum(seqlengths(hits_genome)) - sum(width(peakgr)),
                      sum(width(peakgr[bg_peak_names])),
                      sum(width(peakgr[fg_peak_names]))),
                    names = colnames(n_ov_obs)) / sum(seqlengths(hits_genome)))

    pd_counts <- full_join(as.data.frame(n_ov_obs) |>
                               mutate(motif = "motif") |>
                               pivot_longer(cols = all_of(levels(ov_type)),
                                            names_to = "type",
                                            values_to = "observed"),
                           as.data.frame(n_ov_exp) |>
                               mutate(motif = "motif") |>
                               pivot_longer(cols = all_of(levels(ov_type)),
                                            names_to = "type",
                                            values_to = "expected"),
                           by = join_by(motif, type)) |>
        dplyr::mutate(enr = observed / expected) |>
        pivot_longer(cols = all_of(c("observed", "expected")),
                     names_to = "obsexp", values_to = "val") |>
        dplyr::mutate(type = factor(type, levels = c("non-peak", fgbg_names[2:1]))) |>
        group_by(motif, obsexp) |>
        mutate(tot = sum(val),
               percentage = val / tot) |>
        ungroup()
    pd_counts <- rbind(pd_counts |>
                           filter(obsexp == "observed") |>
                           dplyr::select(motif, type, enr, percentage),
                       pd_counts |>
                           filter(obsexp == "expected" & motif == pd_counts[[1,"motif"]]) |>
                           mutate(motif = "expected") |>
                           dplyr::select(motif, type, enr, percentage)
    ) |> mutate(motif = relevel(factor(motif), "expected"))

    gg_ov_counts <- ggplot(pd_counts, aes(percentage, motif, fill = type)) +
        geom_col(position = position_stack()) +
        scale_x_continuous(labels = scales::label_percent(),
                           expand = expansion(mult = c(0, 0.15))) +
        scale_fill_manual(values = structure(c(gnm_col, fgbg_col[2:1]),
                                             names = levels(pd_counts$type))) +
        geom_text(data = pd_counts |> filter(grepl("foreground", type)),
                  aes(x = 1, label = sprintf("%.1f%%", percentage * 100)),
                  colour = fgbg_col[1], hjust = -0.1, vjust = 0.5,
                  size = fs / ggplot2::.pt,
                  fontface = "bold") +
        labs(x = "Percent of motif sites",
             y = element_blank(),
             fill = "Genomic region: ") +
        theme_cowplot(fs) +
        theme(legend.position = "none") +
        guides(fill = guide_legend(nrow = 3))

    pd_enr <- pd_counts |> filter(motif != "expected") |>
        dplyr::select(c("motif","type","enr")) |> unique()
    levels(pd_enr$type) <- c("non-peak", "backgr.", "foregr.")
    gg_ov_enr <- ggplot(pd_enr, aes(enr, type, fill = type)) +
        geom_col(position = position_dodge2()) +
        scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
        geom_vline(xintercept = 1, linetype = "dotted") +
        scale_fill_manual(values = structure(c(gnm_col, fgbg_col[2:1]),
                                             names = levels(pd_enr$type))) +
        # facet_wrap(~ motif, nrow = 1) +
        labs(x = "Motif site enrichment (obs/exp)",
             y = element_blank(),
             fill = "Genomic region: ") +
        theme_cowplot(fs) +
        theme(legend.position = "bottom") +
        guides(fill = guide_legend(nrow = 3))

    gg_ov_enr_legend <- cowplot::get_legend(gg_ov_enr)
    gg_ov_enr <- gg_ov_enr + theme(legend.position = "none")

    # put panels together
    motif_characterization_plots[[i]] <- cowplot::plot_grid(
        cowplot::plot_grid(gg1_seqlogo, NULL, ncol = 1, rel_heights = c(0.87, 0.13)),
        gg1_kmers,
        gg1_loc,
        cowplot::plot_grid(gg_ov_counts,
                           gg_ov_enr,
                           align = "v", axis = "l",
                           nrow = 2,
                           rel_heights = c(1, 1.2)),
        # align = "h", axis = "tb",
        nrow = 1,
        rel_widths = c(1, 1, 1, 1.4)
    )

    print(motif_characterization_plots[[i]])
}

# add legend
motif_characterization_plots[[length(motif_characterization_plots) + 1]] <-
    cowplot::plot_grid(NULL, NULL, NULL, gg_ov_enr_legend,
                       nrow = 1, rel_widths = c(1, 1, 1, 1.4))
```

## Co-bound peaks for selected transcription factors

```{r co-bound-peaks}
(p_cobound_1 <- create_cobound_peaks_plot(sel_tfs = c("Ams2", "Zas1", "Teb1")))
(p_cobound_2 <- create_cobound_peaks_plot(sel_tfs = c("Esc1", "SPAC3F10.12c")))
```

## Put together

Assemble the panels into Figure 4:

```{r Fig4}
#| fig.width: 11
#| fig.height: 14

cowplot::plot_grid(
    cowplot::plot_grid(
        cowplot::plot_grid(ZnII2Cys6_kmer_1, scale = 0.85),
        hm_ZnII2Cys6_sel_3mers,
        labels = c("A", "B"),
        rel_widths = c(2, 3.2),
        nrow = 1
    ),
    cowplot::plot_grid(
        plotlist = motif_characterization_plots,
        ncol = 1,
        rel_heights = c(rep(1, length(motif_characterization_plots) - 1), 0.4)),
    nrow = 2,
    labels = c("", "C"),
    align = "vh",
    axis = "b",
    rel_heights = c(1.4, 3.2)
)
```

## Supplementary figure

Assemble the panels into Supplementary Figure 4:

```{r suppFig4}
#| fig.width: 4.5
#| fig.height: 6

cowplot::plot_grid(
    p_cobound_2,
    nrow = 1,
    labels = c("A")
)
```

# Session info

<details>
<summary><b>
Session info
</b></summary>
```{r}
sessionInfo()
```
</details>