### 2025-02-20 PomBase_data_export.Rmd
add exports for PomBase
Co-authored-by: Charlotte Soneson <charlottesoneson@gmail.com>
Co-authored-by: Michael Stadler <stadler.michael@gmail.com>

---
title: "PomBase data export"
author: "Charlotte Soneson, Michael Stadler, Merle Skribbe"
date: "`r Sys.Date()`"
output: 
    html_document:
        theme: united
        toc: true
        toc_float: true
        code_folding: hide
        self_contained: true
params:
    rds150: "data/ipms_150_sce.rds"
    rds500: "data/ipms_500_sce.rds"
    peakcsv: "data/fused_peaks_filtered.csv.gz"
    baitclass: "data/ipms_bait_class.txt"
    idmap: "data/id_mapping_table.txt"
    motifrds: "data/motifs_annotated.rds"
    gtffile: "reference/Schizosaccharomyces_pombe.ASM294v2.55.gtf"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = c("png", "cairo_pdf"))
options(scipen = 6)
```

# Goal

This report load, reformats and export data for inclusion in [PomBase](https://www.pombase.org/).

# Parameter values

```{r params}
params
```

# Load required packages and helper functions

```{r load-packages, class.source="fold-show"}
suppressPackageStartupMessages({
    library(SingleCellExperiment)
    library(GenomicFeatures)
    library(universalmotif)
    library(BiocIO)
    library(dplyr)
})

source("params/mapping_functions.R")
```

# Read data

```{r load-data, class.source="fold-show", warning=FALSE}
## Feature ID mapping table
idmap <- read.delim(params$idmap)

## Bait class table
baitclass <- read.delim(params$baitclass)

## IP-MS data
sce150 <- readRDS(params$rds150)
sce500 <- readRDS(params$rds500)

## ChIP-seq peaks
peaks <- read.csv(params$peakcsv, row.names = 1)
peaksgr <- as(peaks, "GenomicRanges")
peakgrNonubi <- as(peaks[peaks$peaktype != "common peaks (ubiquitous)", 1:5],
                   "GenomicRanges")
enrNonubi <- as.matrix(peaks[peaks$peaktype != "common peaks (ubiquitous)",
                             grep("^is_enr_in[.]", colnames(peaks))])
colnames(enrNonubi) <- sub("^is_enr_in[.]", "", colnames(enrNonubi))
enrNonubi <- enrNonubi[, setdiff(colnames(enrNonubi), "Untagged")]

## Transcript database and promoters
txdb <- makeTxDbFromGFF(params$gtffile)
txs <- transcripts(txdb, columns = c("GENEID", "TXNAME"))
colnames(mcols(txs)) <- c("gene_id", "tx_name")
txs$gene_id <- unlist(txs$gene_id, use.names = FALSE)
names(txs) <- txs$tx_name
tss <- resize(txs, width = 1, fix = "start")
prom <- resize(x = tss, width = 1000, fix = "center")

## Motifs
motifs <- readRDS(params$motifrds)
```

# Export data for PomBase

## IP-MS data

Table of proteins pulled-down by each TF at low salt and high salt conditions.

```{r export-ip-ms, class.source="fold-show"}
testL150 <- metadata(sce150)$testres$tests
res150 <- do.call(bind_rows, lapply(names(testL150), function(nm) {
    testL150[[nm]] |>
        dplyr::filter(adj.P.Val < 0.05 & logFC > 1) |>
        dplyr::select(pid, adj.P.Val, logFC) |>
        dplyr::rename(prey = pid) |>
        dplyr::mutate(bait = .getProteinNameFromComparison(nm))
})) |>
    dplyr::filter(bait != prey) |>
    dplyr::filter(!bait %in% baitclass$Gene_name[baitclass$class == "nobait"]) |>
    dplyr::select(bait, prey, adj.P.Val, logFC)

write.table(res150, file = "PomBase/ipms_150mM_pulldowns.csv", row.names = FALSE, 
            col.names = TRUE, sep = ",", quote = FALSE)

testL500 <- metadata(sce500)$testres$tests
res500 <- do.call(bind_rows, lapply(names(testL500), function(nm) {
    testL500[[nm]] |>
        dplyr::filter(adj.P.Val < 0.05 & logFC > 1) |>
        dplyr::select(pid, adj.P.Val, logFC) |>
        dplyr::rename(prey = pid) |>
        dplyr::mutate(bait = .getProteinNameFromComparison(nm))
})) |>
    dplyr::filter(bait != prey) |>
    dplyr::filter(!bait %in% baitclass$Gene_name[baitclass$class == "nobait"]) |>
    dplyr::select(bait, prey, adj.P.Val, logFC)

write.table(res500, file = "PomBase/ipms_500mM_pulldowns.csv", row.names = FALSE, 
            col.names = TRUE, sep = ",", quote = FALSE)
```

## Motifs

Table of identified motif consensus sequences.

```{r export-motifs, class.source="fold-show"}
motiftab <- motifs |>
    dplyr::filter(unique_version_of_motif) |>
    dplyr::select(tf_name, name, consensus) |>
    dplyr::mutate(gene_stable_id = ifelse(
        tf_name %in% idmap$gene_stable_id,
        tf_name,
        idmap$gene_stable_id[match(tolower(tf_name), idmap$gene_name)])) |>
    dplyr::rename(motif_name = name,
                  motif_consensus = consensus) |>
    dplyr::select(tf_name, gene_stable_id, motif_name, motif_consensus)

write.table(motiftab, file = "PomBase/motif_consensus_sequences.csv",
            row.names = FALSE, col.names = TRUE, sep = ",", quote = FALSE)
```

## ChIP-seq data

Table of gene promoters bound by a TF (excluding common ubiquitous peaks).
Promoters are defined as 1kb windows centerd on annotated transcript
start sites. Promoter binding is defined as a peak that overlaps a promoter.

```{r export-bound-promoters, class.source="fold-show"}
prombinding <- do.call(bind_rows, lapply(colnames(enrNonubi), function(tf_name) {
    peakgrSel <- peakgrNonubi[enrNonubi[, tf_name], ]
    ovprom <- overlapsAny(query = prom, subject = peakgrSel)
    data.frame(tf_name = tf_name,
               bound_promoters = paste(unique(prom$gene_id[ovprom]),
                                       collapse = ";"))
}))

# add Zas1 (SPBC1198.04c) promoter to Zas1
# (was missed due to unresolved peak near common-ubiquitous peak)
prombinding[match("Zas1", prombinding$tf_name), "bound_promoters"] <- paste0(
    prombinding[match("Zas1", prombinding$tf_name), "bound_promoters"],
    ";SPBC1198.04c"
)

write.table(prombinding, file = "PomBase/tf_binding_promoters.csv",
            row.names = FALSE, col.names = TRUE, sep = ",", quote = FALSE)
```

## Peaks by type as BED files

```{r export-peaks-bed, class.source="fold-show"}
for (peaktype in unique(peaksgr$peaktype)) {
    peakgrSel <- peaksgr[peaksgr$peaktype == peaktype]
    mcols(peakgrSel) <- NULL
    peaktypeClean <- gsub("[.]", "-", make.names(gsub("[()]", "", peaktype)))
    export(peakgrSel, paste0("PomBase/peaks_", peaktypeClean, ".bed"))
}
```

# Session info

<details>
<summary><b>
Session info
</b></summary>
```{r}
sessionInfo()
```
</details>